---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Test whether any array element along a given axis evaluates to `TRUE`.

@description

# Examples
```{r}
x <- k_array(c(TRUE, FALSE))
k_any(x)
```

```{r}
(x <- k_reshape(c(FALSE, FALSE, FALSE,
                  TRUE, FALSE, FALSE), c(2, 3)))
k_any(x, axis = 1)
k_any(x, axis = 2)
k_any(x, axis = -1)
```

`keepdims = TRUE` outputs a tensor with dimensions reduced to one.
```{r}
k_any(x, keepdims = TRUE)
k_any(x, 1, keepdims = TRUE)
```

@returns
The tensor containing the logical OR reduction over the `axis`.

@param x Input tensor.
@param axis An integer or tuple of integers that represent the axis along
    which a logical OR reduction is performed. The default
    (`axis = NULL`) is to perform a logical OR over all the dimensions
    of the input array. `axis` may be negative, in which case it counts
    for the last to the first axis.
@param keepdims If `TRUE`, axes which are reduced are left in the result as
    dimensions with size one. With this option, the result will
    broadcast correctly against the input array. Defaults to `FALSE`.

@export
@family ops
@seealso
+ <https:/keras.io/keras_core/api/ops/numpy#any-function>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/ops/any>
