---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---

Resize images to size using the specified interpolation method.

@description

# Examples
```{r}
x <- random_uniform(c(2, 4, 4, 3)) # batch of 2 RGB images
y <- k_image_resize(x, c(2, 2))
y$shape
```

```{r}
x <- random_uniform(c(4, 4, 3)) # single RGB image
y <- k_image_resize(x, c(2, 2))
y$shape
```

```{r}
x <- random_uniform(c(2, 3, 4, 4)) # batch of 2 RGB images
y <- k_image_resize(x, c(2, 2), data_format = "channels_first")
y$shape
```

@returns
Resized image or batch of images.

@param image Input image or batch of images. Must be 3D or 4D.
@param size Size of output image in `(height, width)` format.
@param interpolation Interpolation method. Available methods are `"nearest"`,
    `"bilinear"`, and `"bicubic"`. Defaults to `"bilinear"`.
@param antialias Whether to use an antialiasing filter when downsampling an
    image. Defaults to `FALSE`.
@param data_format string, either `"channels_last"` or `"channels_first"`.
    The ordering of the dimensions in the inputs. `"channels_last"`
    corresponds to inputs with shape `(batch, height, width, channels)`
    while `"channels_first"` corresponds to inputs with shape
    `(batch, channels, height, weight)`. It defaults to the
    `image_data_format` value found in your Keras config file at
    `~/.keras/keras.json`. If you never set it, then it will be
    `"channels_last"`.

@export
@family ops
@seealso
+ <https:/keras.io/keras_core/api/ops/image#resize-function>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/ops/image/resize>
