---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the crossentropy metric between the labels and predictions.

@description
Use this crossentropy metric when there are two or more label classes.
It expects labels to be provided as integers. If you want to provide labels
that are one-hot encoded, please use the `metric_categorical_crossentropy()`
metric instead.

There should be `num_classes` floating point values per feature for `y_pred`
and a single floating point value per feature for `y_true`.

# Examples
Standalone usage:

```{r}
m <- metric_sparse_categorical_crossentropy()
m$update_state(c(1, 2),
               rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1)))
m$result()
```

```{r}
m$reset_state()
m$update_state(c(1, 2),
               rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1)),
               sample_weight = c(0.3, 0.7))
m$result()
# 1.6271976
```

Usage with `compile()` API:

```{r, eval = FALSE}
model %>% compile(
    optimizer = 'sgd',
    loss = 'mse',
    metrics = list(metric_sparse_categorical_crossentropy()))
```

@param name
(Optional) string name of the metric instance.

@param dtype
(Optional) data type of the metric result.

@param from_logits
(Optional) Whether output is expected
to be a logits tensor. By default, we consider that output
encodes a probability distribution.

@param axis
(Optional) Defaults to `-1`.
The dimension along which entropy is computed.

@param y_true
Ground truth values.

@param y_pred
The predicted values.

@param ignore_class
Optional integer. The ID of a class to be ignored during
loss computation. This is useful, for example, in segmentation
problems featuring a "void" class (commonly -1 or 255) in
segmentation maps. By default (`ignore_class=NULL`), all classes are
considered.

@param ...
Passed on to the Python callable

@export
@family probabilistic metrics
@family metrics
@seealso
+ <https:/keras.io/api/metrics/probabilistic_metrics#sparsecategoricalcrossentropy-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/metrics/SparseCategoricalCrossentropy>

