---
title: "foRmula keRas"
author: "Pete Mohanty"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: github_document
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE, warning = FALSE)
```


The goal of this document is to introduce `kms` (as in `keras_sequential_model()`), a regression-style function which allows users to call `keras` neural nets with `R` `formula` objects. Formulas are very powerful in R; in the example below, a small tweak in the way the dependent variable is coded explains an additional 20% of out of sample variance. `kms` splits training and test data into sparse matrices and contains the major parameters found in `library(keras)` as inputs (loss function, batch size, number of epochs, etc.). To install my [branch](https://github.com/rdrr1990/keras) of `keras`,

```{r, eval = FALSE}
devtools::install_github("rdrr1990/keras")
```
Let's start with an example using `rtweet` (from `@kearneymw`). The examples here don't provide particularly predictive models so much as show how using `formula` objects can smooth data cleaning and hyperparameter selection.

```{r}
library(rtweet)
rt <- search_tweets("#rstats", n = 5000, include_rts = FALSE)
dim(rt)
summary(rt$retweet_count)
```
Suppose we wanted to predict how many times a tweet with `#rstat` is going to be retweeted. And suppose we wanted to bin the retweent count into five categories (none, 1-10, 11-50, 51-99, and 100 or more). Suppose we believe that the twitter handle and source matters as does day of week and time of day. 
```{r}
library(keras)
breaks <- c(-1, 0, 1, 10, 50, 100, 10000)
out <- kms("cut(retweet_count, breaks) ~ screen_name + source +
            grepl('gg', text) + grepl('tidy', text) + 
            grepl('rstudio', text, ignore.case = TRUE) +
            grepl('cran', text, ignore.case = TRUE) +
            grepl('trump', text, ignore.case = TRUE) +
            weekdays(rt$created_at) + 
            format(rt$created_at, '%d') + 
            format(rt$created_at, '%H')", data = rt)
plot(out$history)
summary(out$model)
out$confusion
out$evaluations
```

Let's say we want to add some data about how many other people are mentioned in each tweet and switch to a (discretized) log scale.

```{r}
rt$Nmentions <- unlist(lapply(rt$mentions_screen_name, 
                              function(x){length(x[[1]]) - is.na(x[[1]])}))

out2 <- kms("floor(log(retweet_count + 1)) ~ Nmentions + screen_name + source +
            grepl('gg', text) + grepl('tidy', text) + 
            grepl('rstudio', text, ignore.case = TRUE) +
            grepl('cran', text, ignore.case = TRUE) +
            grepl('trump', text, ignore.case = TRUE) +
            weekdays(rt$created_at) + 
            format(rt$created_at, '%d') + 
            format(rt$created_at, '%H')", 
            data = rt, Nepochs = 10)
out2$evaluations
out2$confusion
```


Heading in the right direction. Suppose instead we wanted to add who was mentioned.

```{r}
input.formula <- "floor(log(retweet_count + 1)) ~ Nmentions + screen_name + source +
            grepl('gg', text) + grepl('tidy', text) + 
            grepl('rstudio', text, ignore.case = TRUE) + 
            grepl('cran', text, ignore.case = TRUE) +
            grepl('trump', text, ignore.case = TRUE) +
            weekdays(rt$created_at) + format(rt$created_at, '%d') + 
            format(rt$created_at, '%H')"

handles <- names(table(unlist(rt$mentions_screen_name)))

for(i in 1:length(handles)){
  lab <- paste0("mentions_", handles[i])
  rt[[lab]] <- grepl(handles[i], rt$mentions_screen_name)
  input.formula <- paste(input.formula, "+", lab)
}

out3 <- kms(input.formula, data = rt, Nepochs = 10)
out3$evaluations
out3$confusion
```
Marginal improvement but the model is still clearly overpredicting the modal outcome (zero retweets) and struggling to forecast the rare, popular tweets. Maybe the model needs more layers. 

```{r}
out4 <- kms(input.formula, data = rt, 
            layers = list(units = c(405, 135, 45, 15, NA), 
                         activation = c("softmax", "relu", "relu", "relu", "softmax"), 
                         dropout = c(0.7, 0.6, 0.5, 0.4, NA)),
            Nepochs = 6)
out4$evaluations
out4$confusion
```
Suppose we wanted to see if the estimates were stable across 10 test/train splits.

```{r}
est <- list()
accuracy <- c()
for(i in 1:10){
  est[[paste0("seed", i)]] <- kms(input.formula, rt, seed = i,
            layers = list(units = c(405, 135, 45, 15, NA), 
                         activation = c("softmax", "relu", "relu", "relu", "softmax"), 
                         dropout = c(0.7, 0.6, 0.5, 0.4, NA)),
            Nepochs = 10)
  accuracy[i] <- est[[paste0("seed", i)]][["evaluations"]][["acc"]]
}
accuracy
```

Hmmm... Maybe Model 3 is the closest ... Or maybe we just need more data :)

Though `kms` contains a number of parameters, the goal is not to replace all the vast customizability that `keras` offers. Rather, like `qplot` in the `ggplot` library, `kms` offers convenience for common scenarios. Or, perhaps better, like `MCMCpack` or `rstan` do for Bayesian MCMC, `kms` aims to introduce users familiar with regression in R to neural nets without steep scripting stumbling blocks. Next on the to-do list is to add support for accepting keras models as a parameter. Suggestions are more than welcome!