---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the crossentropy loss between the labels and predictions.

@description
Use this crossentropy loss function when there are two or more label
classes. We expect labels to be provided in a `one_hot` representation. If
you want to provide labels as integers, please use
`SparseCategoricalCrossentropy` loss. There should be `num_classes` floating
point values per feature, i.e., the shape of both `y_pred` and `y_true` are
`[batch_size, num_classes]`.

# Examples
```{r}
y_true <- rbind(c(0, 1, 0), c(0, 0, 1))
y_pred <- rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1))
loss <- loss_categorical_crossentropy(y_true, y_pred)
loss
```
Standalone usage:

```{r}
y_true <- rbind(c(0, 1, 0), c(0, 0, 1))
y_pred <- rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1))
# Using 'auto'/'sum_over_batch_size' reduction type.
cce <- loss_categorical_crossentropy()
cce(y_true, y_pred)
```

```{r}
# Calling with 'sample_weight'.
cce(y_true, y_pred, sample_weight = k_array(c(0.3, 0.7)))
```

```{r}
# Using 'sum' reduction type.
cce <- loss_categorical_crossentropy(reduction = "sum")
cce(y_true, y_pred)
```

```{r}
# Using 'none' reduction type.
cce <- loss_categorical_crossentropy(reduction = NULL)
cce(y_true, y_pred)
```

Usage with the `compile()` API:

```{r, eval = FALSE}
model %>% compile(optimizer = 'sgd',
              loss=loss_categorical_crossentropy())
```

@returns
Categorical crossentropy loss value.

@param from_logits
Whether `y_pred` is expected to be a logits tensor. By
default, we assume that `y_pred` encodes a probability distribution.

@param label_smoothing
Float in `[0, 1].` When > 0, label values are smoothed,
meaning the confidence on label values are relaxed. For example, if
`0.1`, use `0.1 / num_classes` for non-target labels and
`0.9 + 0.1 / num_classes` for target labels.

@param axis
The axis along which to compute crossentropy (the features
axis). Defaults to `-1`.

@param reduction
Type of reduction to apply to the loss. In almost all cases
this should be `"sum_over_batch_size"`.
Supported options are `"sum"`, `"sum_over_batch_size"` or `NULL`.

@param name
Optional name for the loss instance.

@param y_true
Tensor of one-hot true targets.

@param y_pred
Tensor of predicted targets.

@param ...
Passed on to the Python callable

@export
@family losses
@seealso
+ <https:/keras.io/api/losses/probabilistic_losses#categoricalcrossentropy-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/losses/CategoricalCrossentropy>

