---
title: "Writing `Layer` and `Model` objects from scratch."
author: "[fchollet](https://twitter.com/fchollet), [t-kalinowski](https://github.com/t-kalinowski)"
output: rmarkdown::html_vignette
date created: 2019/03/01
last modified: 2020/04/13
description: Guide to writing `Layer` and `Model` objects from scratch.
vignette: >
  %\VignetteIndexEntry{Writing `Layer` and `Model` objects from scratch.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Setup
```{r, echo=FALSE}
Sys.setenv(TF_CPP_MIN_LOG_LEVEL = 2)
knitr::opts_chunk$set(comment = NA, eval = FALSE)
```

```{r}
library(magrittr)
library(tensorflow)
library(tfdatasets)
library(keras)

tf_version()
```


## The `Layer` class: a combination of state (weights) and some computation

One of the central abstractions in Keras is the `Layer` class. A layer
encapsulates both a state (the layer's "weights") and a transformation from
inputs to outputs (a "call", the layer's forward pass).

Here's a densely-connected layer. It has a state: the variables `w` and `b`.
```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32, input_dim = 32) {
    super$initialize()
    w_init <- tf$random_normal_initializer()
    self$w <- tf$Variable(
      initial_value = w_init(
        shape = shape(input_dim, units),
        dtype = "float32"
      ),
      trainable = TRUE
    )
    b_init <- tf$zeros_initializer()
    self$b <- tf$Variable(
      initial_value = b_init(shape = shape(units), dtype = "float32"),
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }
}
```

You would use a layer by calling it on some tensor input(s), much like a regular
function.

```{r}
x <- tf$ones(shape(2, 2))
linear_layer <- Linear(4, 2)
y <- linear_layer(x)
print(y)
```

`Linear` behaves similarly to a layer present in the python interface to keras
(e.g., `keras$layers$Dense`).

However, one additional step is needed to make it behave like the builtin layers
present in the keras R package (e.g., `layer_dense()`).

Keras layers in R are designed to compose nicely with the pipe operator
(`%>%`), so that the layer instance is conveniently created on demand when an
existing model or tensor is piped in. In order to make a custom layer similarly
compose nicely with the pipe, you can call `create_layer_wrapper()` on the layer
class constructor.

```{r}
layer_linear <- create_layer_wrapper(Linear)
```

Now `layer_linear` is a layer constructor that composes nicely with `%>%`, just like the built-in layers:

```{r}
model <- keras_model_sequential() %>%
  layer_linear(4, 2)

model(k_ones(c(2, 2)))

model
```

Because the pattern above is so common, there is a convenience function that
combines the steps of subclassing `keras$layers$Layer` and calling
`create_layer_wrapper` on the output: the `Layer` function. The `layer_linear`
defined below is identical to the `layer_linear` defined above.

```{r}
layer_linear <- Layer(
  "Linear",
  initialize =  function(units = 32, input_dim = 32) {
    super$initialize()
    w_init <- tf$random_normal_initializer()
    self$w <- tf$Variable(initial_value = w_init(shape = shape(input_dim, units),
                                                 dtype = "float32"),
                          trainable = TRUE)
    b_init <- tf$zeros_initializer()
    self$b <- tf$Variable(initial_value = b_init(shape = shape(units),
                                                 dtype = "float32"),
                          trainable = TRUE)
  },

  call = function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }
)
```

For the remainder of this vignette we'll be using the `%py_class%` constructor. However,
in your own code feel free to use `create_layer_wrapper` and/or `Layer` if you prefer.

Note that the weights `w` and `b` are automatically tracked by the layer upon being set as
layer attributes:

```{r}
stopifnot(all.equal(
  linear_layer$weights,
  list(linear_layer$w, linear_layer$b)
))
```

You also have access to a quicker shortcut for adding a weight to a layer:
the `add_weight()` method:

```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32, input_dim = 32) {
    super$initialize()
    w_init <- tf$random_normal_initializer()
    self$w <- self$add_weight(
      shape = shape(input_dim, units),
      initializer = "random_normal",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(units),
      initializer = "zeros",
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }
}

x <- tf$ones(shape(2, 2))
linear_layer <- Linear(4, 2)
y <- linear_layer(x)
print(y)
```

## Layers can have non-trainable weights

Besides trainable weights, you can add non-trainable weights to a layer as
well. Such weights are meant not to be taken into account during
backpropagation, when you are training the layer.

Here's how to add and use a non-trainable weight:

```{r}
ComputeSum(keras$layers$Layer) %py_class% {
  initialize <- function(input_dim) {
    super$initialize()
    self$total <- tf$Variable(
      initial_value = tf$zeros(shape(input_dim)),
      trainable = FALSE
    )
  }

  call <- function(inputs) {
    self$total$assign_add(tf$reduce_sum(inputs, axis = 0L))
    self$total
  }
}

x <- tf$ones(shape(2, 2))
my_sum <- ComputeSum(2)
y <- my_sum(x)
print(as.numeric(y))
y <- my_sum(x)
print(as.numeric(y))
```

It's part of `layer$weights`, but it gets categorized as a non-trainable weight:

```{r}
cat("weights:", length(my_sum$weights), "\n")
cat("non-trainable weights:", length(my_sum$non_trainable_weights), "\n")

# It's not included in the trainable weights:
cat("trainable_weights:", my_sum$trainable_weights, "\n")
```

## Best practice: deferring weight creation until the shape of the inputs is known

Our `Linear` layer above took an `input_dim `argument that was used to compute
the shape of the weights `w` and `b` in `initialize()`:


```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32, input_dim = 32) {
    super$initialize()
    self$w <- self$add_weight(
      shape = shape(input_dim, units),
      initializer = "random_normal",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(units),
      initializer = "zeros",
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }
}
```

In many cases, you may not know in advance the size of your inputs, and you
would like to lazily create weights when that value becomes known, some time
after instantiating the layer.

In the Keras API, we recommend creating layer weights in the `build(self,
inputs_shape)` method of your layer. Like this:

```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32) {
    super$initialize()
    self$units <- units
  }

  build <- function(input_shape) {
    self$w <- self$add_weight(
      shape = shape(tail(input_shape, 1), self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }
}
```

The `build()` method of your layer will automatically run the first time
your layer instance is called. You now have a layer that can handle an arbitrary number of input features:

```{r}
# At instantiation, we don't know on what inputs this is going to get called
linear_layer <- Linear(32)

# The layer's weights are created dynamically the first time the layer is called
y <- linear_layer(x)
```



## Layers are recursively composable

If you assign a Layer instance as an attribute of another Layer, the outer layer
will start tracking the weights of the inner layer.

We recommend creating such sublayers in the `initialize()` method (since the
sublayers will typically have a build method, they will be built when the
outer layer gets built).



```{r}
# Let's assume we are reusing the Linear class
# with a `build` method that we defined above.
MLPBlock(keras$layers$Layer) %py_class% {
  initialize <- function() {
    super$initialize()
    self$linear_1 <- Linear(32)
    self$linear_2 <- Linear(32)
    self$linear_3 <- Linear(1)
  }

  call <- function(inputs) {
    x <- self$linear_1(inputs)
    x <- tf$nn$relu(x)
    x <- self$linear_2(x)
    x <- tf$nn$relu(x)
    self$linear_3(x)
  }
}

mlp <- MLPBlock()
y <- mlp(tf$ones(shape = shape(3, 64))) # The first call to the `mlp` will create the weights
cat("weights:", length(mlp$weights), "\n")
cat("trainable weights:", length(mlp$trainable_weights), "\n")
```

## The `add_loss()` method

When writing the `call()` method of a layer, you can create loss tensors that
you will want to use later, when writing your training loop. This is doable by
calling `self$add_loss(value)`:


```{r}
# A layer that creates an activity regularization loss
ActivityRegularizationLayer(keras$layers$Layer) %py_class% {
  initialize <- function(rate = 1e-2) {
    super$initialize()
    self$rate <- rate
  }

  call <- function(inputs) {
    self$add_loss(self$rate * tf$reduce_sum(inputs))
    inputs
  }
}
```

These losses (including those created by any inner layer) can be retrieved via
`layer$losses`. This property is reset at the start of every `call()` to
the top-level layer, so that `layer$losses` always contains the loss values
created during the last forward pass.


```{r}
OuterLayer(keras$layers$Layer) %py_class% {
  initialize <- function() {
    super$initialize()
    self$activity_reg <- ActivityRegularizationLayer(1e-2)
  }
  call <- function(inputs) {
    self$activity_reg(inputs)
  }
}

layer <- OuterLayer()
stopifnot(length(layer$losses) == 0) # No losses yet since the layer has never been called

layer(tf$zeros(shape(1, 1))) |> invisible()
stopifnot(length(layer$losses) == 1) # We created one loss value

# `layer$losses` gets reset at the start of each call()
layer(tf$zeros(shape(1, 1))) |> invisible()
stopifnot(length(layer$losses) == 1) # This is the loss created during the call above
```

In addition, the `loss` property also contains regularization losses created
for the weights of any inner layer:

```{r}
OuterLayerWithKernelRegularizer(keras$layers$Layer) %py_class% {
  initialize <- function() {
    super$initialize()
    self$dense <- layer_dense(units = 32, kernel_regularizer = regularizer_l2(1e-3))
  }
  call <- function(inputs) {
    self$dense(inputs)
  }
}

layer <- OuterLayerWithKernelRegularizer()
layer(tf$zeros(shape(1, 1))) |> invisible()

# This is `1e-3 * sum(layer$dense$kernel ** 2)`,
# created by the `kernel_regularizer` above.
print(layer$losses)
```

These losses are meant to be taken into account when writing training loops,
like this:

```{r, eval=FALSE}
# Instantiate an optimizer.
optimizer <- optimizer_sgd(learning_rate = 1e-3)
loss_fn <- loss_sparse_categorical_crossentropy(from_logits = TRUE)

# Iterate over the batches of a dataset.
dataset_iterator <- reticulate::as_iterator(train_dataset)
while(!is.null(batch <- iter_next(dataset_iterator))) {
  c(x_batch_train, y_batch_train) %<-% batch
  with(tf$GradientTape() %as% tape, {
    logits <- layer(x_batch_train) # Logits for this minibatch
    # Loss value for this minibatch
    loss_value <- loss_fn(y_batch_train, logits)
    # Add extra losses created during this forward pass:
    loss_value <- loss_value + sum(model$losses)
  })
  grads <- tape$gradient(loss_value, model$trainable_weights)
  optimizer$apply_gradients(
    purrr::transpose(list(grads, model$trainable_weights)))
}
```

For a detailed guide about writing training loops, see the
[guide to writing a training loop from scratch](/guides/writing_a_training_loop_from_scratch/).

These losses also work seamlessly with `fit()` (they get automatically summed
and added to the main loss, if any):

```{r}
input <- layer_input(shape(3))
output <- input %>% layer_activity_regularization()
# output <- ActivityRegularizationLayer()(input)
model <- keras_model(input, output)

# If there is a loss passed in `compile`, the regularization
# losses get added to it
model %>% compile(optimizer = "adam", loss = "mse")
model %>% fit(k_random_uniform(c(2, 3)),
  k_random_uniform(c(2, 3)),
  epochs = 1, verbose = FALSE
)

# It's also possible not to pass any loss in `compile`,
# since the model already has a loss to minimize, via the `add_loss`
# call during the forward pass!
model %>% compile(optimizer = "adam")
model %>% fit(k_random_uniform(c(2, 3)),
  k_random_uniform(c(2, 3)),
  epochs = 1, verbose = FALSE
)
```

## The `add_metric()` method

Similarly to `add_loss()`, layers also have an `add_metric()` method
for tracking the moving average of a quantity during training.

Consider the following layer: a "logistic endpoint" layer.
It takes as inputs predictions and targets, it computes a loss which it tracks
via `add_loss()`, and it computes an accuracy scalar, which it tracks via
`add_metric()`.

```{r}
LogisticEndpoint(keras$layers$Layer) %py_class% {
  initialize <- function(name = NULL) {
    super$initialize(name = name)
    self$loss_fn <- loss_binary_crossentropy(from_logits = TRUE)
    self$accuracy_fn <- metric_binary_accuracy()
  }

  call <- function(targets, logits, sample_weights = NULL) {
    # Compute the training-time loss value and add it
    # to the layer using `self$add_loss()`.
    loss <- self$loss_fn(targets, logits, sample_weights)
    self$add_loss(loss)

    # Log accuracy as a metric and add it
    # to the layer using `self.add_metric()`.
    acc <- self$accuracy_fn(targets, logits, sample_weights)
    self$add_metric(acc, name = "accuracy")

    # Return the inference-time prediction tensor (for `.predict()`).
    tf$nn$softmax(logits)
  }
}
```



Metrics tracked in this way are accessible via `layer$metrics`:

```{r}
layer <- LogisticEndpoint()

targets <- tf$ones(shape(2, 2))
logits <- tf$ones(shape(2, 2))
y <- layer(targets, logits)

cat("layer$metrics: ")
str(layer$metrics)
cat("current accuracy value:", as.numeric(layer$metrics[[1]]$result()), "\n")
```

Just like for `add_loss()`, these metrics are tracked by `fit()`:

```{r}
inputs <- layer_input(shape(3), name = "inputs")
targets <- layer_input(shape(10), name = "targets")
logits <- inputs %>% layer_dense(10)
predictions <- LogisticEndpoint(name = "predictions")(logits, targets)

model <- keras_model(inputs = list(inputs, targets), outputs = predictions)
model %>% compile(optimizer = "adam")

data <- list(
  inputs = k_random_uniform(c(3, 3)),
  targets = k_random_uniform(c(3, 10))
)

model %>% fit(data, epochs = 1, verbose = FALSE)
```

## You can optionally enable serialization on your layers

If you need your custom layers to be serializable as part of a
[Functional model](/guides/functional_api/), you can optionally implement a `get_config()`
method:

```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32) {
    super$initialize()
    self$units <- units
  }

  build <- function(input_shape) {
    self$w <- self$add_weight(
      shape = shape(tail(input_shape, 1), self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }

  get_config <- function() {
    list(units = self$units)
  }
}


# Now you can recreate the layer from its config:
layer <- Linear(64)
config <- layer$get_config()
print(config)
new_layer <- Linear$from_config(config)
```

Note that the `initialize()` method of the base `Layer` class takes some additional named
arguments, in particular a `name` and a `dtype`. It's good practice to pass
these arguments to the parent class in `initialize()` and to include them in the
layer config:

```{r}
Linear(keras$layers$Layer) %py_class% {
  initialize <- function(units = 32, ...) {
    super$initialize(...)
    self$units <- units
  }

  build <- function(input_shape) {
    self$w <- self$add_weight(
      shape = shape(tail(input_shape, 1), self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(self$units),
      initializer = "random_normal",
      trainable = TRUE
    )
  }

  call <- function(inputs) {
    tf$matmul(inputs, self$w) + self$b
  }

  get_config <- function() {
    config <- super$get_config()
    config$units <- self$units
    config
  }
}


layer <- Linear(64)
config <- layer$get_config()
str(config)
new_layer <- Linear$from_config(config)
```

If you need more flexibility when deserializing the layer from its config, you
can also override the `from_config()` class method. This is the base
implementation of `from_config()`:

```{r}
from_config <- function(cls, config) do.call(cls, config)
```


To learn more about serialization and saving, see the complete
[guide to saving and serializing models](/guides/serialization_and_saving/).


## Privileged `training` argument in the `call()` method

Some layers, in particular the `BatchNormalization` layer and the `Dropout`
layer, have different behaviors during training and inference. For such
layers, it is standard practice to expose a `training` (boolean) argument in
the `call()` method.

By exposing this argument in `call()`, you enable the built-in training and
evaluation loops (e.g. `fit()`) to correctly use the layer in training and
inference. Note, the default of `NULL` means that the training parameter will be
inferred by keras from the training context (e.g., it will be `TRUE` if called
from `fit()`, `FALSE` if called from `predict()`)

```{r}
CustomDropout(keras$layers$Layer) %py_class% {
  initialize <- function(rate, ...) {
    super$initialize(...)
    self$rate <- rate
  }
  call <- function(inputs, training = NULL) {
    if (isTRUE(training)) {
      return(tf$nn$dropout(inputs, rate = self$rate))
    }
    inputs
  }
}
```

## Privileged `mask` argument in the `call()` method

The other privileged argument supported by `call()` is the `mask` argument.

You will find it in all Keras RNN layers. A mask is a boolean tensor (one
boolean value per timestep in the input) used to skip certain input timesteps
when processing timeseries data.

Keras will automatically pass the correct `mask` argument to `call()` for
layers that support it, when a mask is generated by a prior layer.
Mask-generating layers are the `Embedding`
layer configured with `mask_zero=True`, and the `Masking` layer.

To learn more about masking and how to write masking-enabled layers, please
check out the guide
["understanding padding and masking"](/guides/understanding_masking_and_padding/).


## The `Model` class

In general, you will use the `Layer` class to define inner computation blocks,
and will use the `Model` class to define the outer model -- the object you
will train.

For instance, in a ResNet50 model, you would have several ResNet blocks
subclassing `Layer`, and a single `Model` encompassing the entire ResNet50
network.

The `Model` class has the same API as `Layer`, with the following differences:

- It has support for built-in training, evaluation, and prediction methods
  (`fit()`, `evaluate()`, `predict()`).
- It exposes the list of its inner layers, via the `model$layers` property.
- It exposes saving and serialization APIs (`save_model_tf()`, `save_model_weights_tf()`, ...)

Effectively, the `Layer` class corresponds to what we refer to in the
literature as a "layer" (as in "convolution layer" or "recurrent layer") or as
a "block" (as in "ResNet block" or "Inception block").

Meanwhile, the `Model` class corresponds to what is referred to in the
literature as a "model" (as in "deep learning model") or as a "network" (as in
"deep neural network").

So if you're wondering, "should I use the `Layer` class or the `Model` class?",
ask yourself: will I need to call `fit()` on it? Will I need to call `save()`
on it? If so, go with `Model`. If not (either because your class is just a block
in a bigger system, or because you are writing training & saving code yourself),
use `Layer`.

For instance, we could take our mini-resnet example above, and use it to build
a `Model` that we could train with `fit()`, and that we could save with
`save_model_weights_tf()`:

```{r, eval = FALSE}
ResNet(keras$Model) %py_class% {
  initialize <- function(num_classes = 1000) {
    super$initialize()
    self$block_1 <- ResNetBlock()
    self$block_2 <- ResNetBlock()
    self$global_pool <- layer_global_average_pooling_2d()
    self$classifier <- layer_dense(units = num_classes)
  }

  call <- function(inputs) {
    x <- self$block_1(inputs)
    x <- self$block_2(x)
    x <- self$global_pool(x)
    self$classifier(x)
  }
}


resnet <- ResNet()
dataset <- ...
resnet %>% fit(dataset, epochs = 10)
resnet %>% save_model_tf(filepath)
```


## Putting it all together: an end-to-end example

Here's what you've learned so far:

- A `Layer` encapsulates a state (created in `initialize()` or `build()`), and some
computation (defined in `call()`).
- Layers can be recursively nested to create new, bigger computation blocks.
- Layers can create and track losses (typically regularization losses) as well
as metrics, via `add_loss()` and `add_metric()`
- The outer container, the thing you want to train, is a `Model`. A `Model` is
just like a `Layer`, but with added training and serialization utilities.

Let's put all of these things together into an end-to-end example: we're going
to implement a Variational AutoEncoder (VAE). We'll train it on MNIST digits.

Our VAE will be a subclass of `Model`, built as a nested composition of layers
that subclass `Layer`. It will feature a regularization loss (KL divergence).

```{r}
Sampling(keras$layers$Layer) %py_class% {
  call <- function(inputs) {
    c(z_mean, z_log_var) %<-% inputs
    batch <- tf$shape(z_mean)[1]
    dim <- tf$shape(z_mean)[2]
    epsilon <- k_random_normal(shape = c(batch, dim))
    z_mean + exp(0.5 * z_log_var) * epsilon
  }
}


Encoder(keras$layers$Layer) %py_class% {
  "Maps MNIST digits to a triplet (z_mean, z_log_var, z)."

  initialize <- function(latent_dim = 32, intermediate_dim = 64, name = "encoder", ...) {
    super$initialize(name = name, ...)
    self$dense_proj <- layer_dense(units = intermediate_dim, activation = "relu")
    self$dense_mean <- layer_dense(units = latent_dim)
    self$dense_log_var <- layer_dense(units = latent_dim)
    self$sampling <- Sampling()
  }

  call <- function(inputs) {
    x <- self$dense_proj(inputs)
    z_mean <- self$dense_mean(x)
    z_log_var <- self$dense_log_var(x)
    z <- self$sampling(c(z_mean, z_log_var))
    list(z_mean, z_log_var, z)
  }
}


Decoder(keras$layers$Layer) %py_class% {
  "Converts z, the encoded digit vector, back into a readable digit."

  initialize <- function(original_dim, intermediate_dim = 64, name = "decoder", ...) {
    super$initialize(name = name, ...)
    self$dense_proj <- layer_dense(units = intermediate_dim, activation = "relu")
    self$dense_output <- layer_dense(units = original_dim, activation = "sigmoid")
  }

  call <- function(inputs) {
    x <- self$dense_proj(inputs)
    self$dense_output(x)
  }
}


VariationalAutoEncoder(keras$Model) %py_class% {
  "Combines the encoder and decoder into an end-to-end model for training."

  initialize <- function(original_dim, intermediate_dim = 64, latent_dim = 32,
                         name = "autoencoder", ...) {
    super$initialize(name = name, ...)
    self$original_dim <- original_dim
    self$encoder <- Encoder(
      latent_dim = latent_dim,
      intermediate_dim = intermediate_dim
    )
    self$decoder <- Decoder(original_dim, intermediate_dim = intermediate_dim)
  }

  call <- function(inputs) {
    c(z_mean, z_log_var, z) %<-% self$encoder(inputs)
    reconstructed <- self$decoder(z)
    # Add KL divergence regularization loss.
    kl_loss <- -0.5 * tf$reduce_mean(z_log_var - tf$square(z_mean) - tf$exp(z_log_var) + 1)
    self$add_loss(kl_loss)
    reconstructed
  }
}
```

Let's write a simple training loop on MNIST:

```{r}
library(tfautograph)
library(tfdatasets)


original_dim <- 784
vae <- VariationalAutoEncoder(original_dim, 64, 32)

optimizer <- optimizer_adam(learning_rate = 1e-3)
mse_loss_fn <- loss_mean_squared_error()

loss_metric <- metric_mean()

x_train <- dataset_mnist()$train$x %>%
  array_reshape(c(60000, 784)) %>%
  `/`(255)

train_dataset <- tensor_slices_dataset(x_train) %>%
  dataset_shuffle(buffer_size = 1024) %>%
  dataset_batch(64)

epochs <- 2

# Iterate over epochs.
for (epoch in seq(epochs)) {
  cat(sprintf("Start of epoch %d\n", epoch))

  # Iterate over the batches of the dataset.
  # autograph lets you use tfdatasets in `for` and `while`
  autograph({
    step <- 0
    for (x_batch_train in train_dataset) {
      with(tf$GradientTape() %as% tape, {
        ## Note: we're four opaque contexts deep here (for, autograph, for,
        ## with), When in doubt about the objects or methods that are available
        ## (e.g., what is `tape` here?), remember you can always drop into a
        ## debugger right here:
        # browser()

        reconstructed <- vae(x_batch_train)
        # Compute reconstruction loss
        loss <- mse_loss_fn(x_batch_train, reconstructed)

        loss %<>% add(vae$losses[[1]]) # Add KLD regularization loss
      })
      grads <- tape$gradient(loss, vae$trainable_weights)
      optimizer$apply_gradients(
        purrr::transpose(list(grads, vae$trainable_weights)))

      loss_metric(loss)

      step %<>% add(1)
      if (step %% 100 == 0) {
        cat(sprintf("step %d: mean loss = %.4f\n", step, loss_metric$result()))
      }
    }
  })
}
```

Note that since the VAE is subclassing `Model`, it features built-in training
loops. So you could also have trained it like this:

```{r}
vae <- VariationalAutoEncoder(784, 64, 32)

optimizer <- optimizer_adam(learning_rate = 1e-3)

vae %>% compile(optimizer, loss = loss_mean_squared_error())
vae %>% fit(x_train, x_train, epochs = 2, batch_size = 64)
```

## Beyond object-oriented development: the Functional API

If you prefer a less object-oriented way of programming, you can also
build models using the [Functional API](/guides/functional_api/). Importantly,
choosing one style or another does not prevent you from leveraging components
written in the other style: you can always mix-and-match.

For instance, the Functional API example below reuses the same `Sampling` layer
we defined in the example above:

```{r}
original_dim <- 784
intermediate_dim <- 64
latent_dim <- 32

# Define encoder model.
original_inputs <- layer_input(shape = original_dim, name = "encoder_input")
x <- layer_dense(units = intermediate_dim, activation = "relu")(original_inputs)
z_mean <- layer_dense(units = latent_dim, name = "z_mean")(x)
z_log_var <- layer_dense(units = latent_dim, name = "z_log_var")(x)
z <- Sampling()(list(z_mean, z_log_var))
encoder <- keras_model(inputs = original_inputs, outputs = z, name = "encoder")

# Define decoder model.
latent_inputs <- layer_input(shape = latent_dim, name = "z_sampling")
x <- layer_dense(units = intermediate_dim, activation = "relu")(latent_inputs)
outputs <- layer_dense(units = original_dim, activation = "sigmoid")(x)
decoder <- keras_model(inputs = latent_inputs, outputs = outputs, name = "decoder")

# Define VAE model.
outputs <- decoder(z)
vae <- keras_model(inputs = original_inputs, outputs = outputs, name = "vae")

# Add KL divergence regularization loss.
kl_loss <- -0.5 * tf$reduce_mean(z_log_var - tf$square(z_mean) - tf$exp(z_log_var) + 1)
vae$add_loss(kl_loss)

# Train.
optimizer <- keras$optimizers$Adam(learning_rate = 1e-3)
vae %>% compile(optimizer, loss = loss_mean_squared_error())
vae %>% fit(x_train, x_train, epochs = 3, batch_size = 64)
```

For more information, make sure to read the [Functional API guide](/guides/functional_api/).


## Defining custom layers and models in an R package

Unfortunately you can't use anything that creates references to python objects (e.g., `%py_class%`), at the top-level of an R package.

When you build an R package, all the R files in the `R/` directory get sourced
in an R environment (the package namespace), and then that environment is saved
as part of the package bundle. Loading the package means restoring the saved R
environment. This means that the R code only gets sourced once, at build time.
If you create references to external objects (e.g., python objects) at package
build time, they will be NULL pointers when the package is loaded, because the
external objects they pointed to at build time no longer exist at load time.

The solution is to delay creating references to python objects until run time. There are a few ways to do this:


+ `Layer()` is lazy about initializing the python reference, so it is safe to use in a package.

+ `%py_class%` is itself a wrapper around `r_to_py(R6Class(...))`. To avoid
calling `r_to_py` at build time, you can define your custom layers using
`create_layer_wrapper(R6Class(...))`. `create_layer_wrapper` will delay calling
`r_to_py` on the R6Class until the first time a layer instance is created, so it
is safe to use in package. Here is an example for how the `layer_linear` could
be safely defined in an R package with this approach.

```{r}
Linear <- R6::R6Class(
  classname = "Linear",
  inherit = keras$layers$Layer,
  public = list(
    initialize = function(units = 32) {
      super$initialize()
      self$units <- units
    },
    build = function(input_shape) {
      self$w <- self$add_weight(
        shape = shape(tail(input_shape, 1), self$units),
        initializer = "random_normal",
        trainable = TRUE
      )
      self$b <- self$add_weight(
        shape = shape(self$units),
        initializer = "random_normal",
        trainable = TRUE
      )
    },
    call = function(inputs) {
      tf$matmul(inputs, self$w) + self$b
    }
  )
)

#' @export
layer_linear <- create_layer_wrapper(Linear)
```

+ If the above approach doesn't satisfy your needs, you can still use
`%py_class%` so long as it is called only after the package is loaded and the
python session is initialized.

If you're writing an R package that uses keras and reticulate,
[this article](https://rstudio.github.io/reticulate/articles/package.html) might be helpful.

## Summary

In this guide you learned about creating custom layers and models in keras.

+  The constructors available: `%py_class%`, `create_layer_wrapper()`, `R6Class()`, and `Layer()`.
+  What methods to you might want to define to your model: `initialize()`, `build()`, `call()`, and `get_config()`.
+  What convenience methods are available when you subclass `keras$layers$Layer`: `add_weight()`, `add_loss()`, and `add_metric()`
