---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Average pooling for temporal data.

@description
Downsamples the input representation by taking the average value over the
window defined by `pool_size`. The window is shifted by `strides`.  The
resulting output when using "valid" padding option has a shape of:
`output_shape = (input_shape - pool_size + 1) / strides)`

The resulting output shape when using the "same" padding option is:
`output_shape = input_shape / strides`

# Input Shape
- If `data_format="channels_last"`:
    3D tensor with shape `(batch_size, steps, features)`.
- If `data_format="channels_first"`:
    3D tensor with shape `(batch_size, features, steps)`.

# Output Shape
- If `data_format="channels_last"`:
    3D tensor with shape `(batch_size, downsampled_steps, features)`.
- If `data_format="channels_first"`:
    3D tensor with shape `(batch_size, features, downsampled_steps)`.

# Examples
`strides=1` and `padding="valid"`:

```{r}
x <- k_array(c(1., 2., 3., 4., 5.)) |> k_reshape(c(1, 5, 1))
output <- x |>
  layer_average_pooling_1d(pool_size = 2,
                           strides = 1,
                           padding = "valid")
output
```

`strides=2` and `padding="valid"`:

```{r}
x <- k_array(c(1., 2., 3., 4., 5.)) |> k_reshape(c(1, 5, 1))
output <- x |>
  layer_average_pooling_1d(pool_size = 2,
                           strides = 2,
                           padding = "valid")
output
```

`strides=1` and `padding="same"`:

```{r}
x <- k_array(c(1., 2., 3., 4., 5.)) |> k_reshape(c(1, 5, 1))
output <- x |>
  layer_average_pooling_1d(pool_size = 2,
                           strides = 1,
                           padding = "same")
output
```

@param pool_size
int, size of the max pooling window.

@param strides
int or `NULL`. Specifies how much the pooling window moves
for each pooling step. If `NULL`, it will default to `pool_size`.

@param padding
string, either `"valid"` or `"same"` (case-insensitive).
`"valid"` means no padding. `"same"` results in padding evenly to
the left/right or up/down of the input such that output has the same
height/width dimension as the input.

@param data_format
string, either `"channels_last"` or `"channels_first"`.
The ordering of the dimensions in the inputs. `"channels_last"`
corresponds to inputs with shape `(batch, steps, features)`
while `"channels_first"` corresponds to inputs with shape
`(batch, features, steps)`. It defaults to the `image_data_format`
value found in your Keras config file at `~/.keras/keras.json`.
If you never set it, then it will be `"channels_last"`.

@param object
Object to compose the layer with. A tensor, array, or sequential model.

@param name
String, name for the object

@param ...
Passed on to the Python callable

@export
@family pooling layers
@family layers
@seealso
+ <https:/keras.io/api/layers/pooling_layers/average_pooling1d#averagepooling1d-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/layers/AveragePooling1D>
