---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes element-wise dot product of two tensors.

@description
It takes a list of inputs of size 2, and the axes
corresponding to each input along with the dot product
is to be performed.

Let's say `x` and `y` are the two input tensors with shapes
`(2, 3, 5)` and `(2, 10, 3)`. The batch dimension should be
of same size for both the inputs, and `axes` should correspond
to the dimensions that have the same size in the corresponding
inputs. e.g. with `axes = c(1, 2)`, the dot product of `x`, and `y`
will result in a tensor with shape `(2, 5, 10)`

# Examples
```{r}
x <- k_reshape(0:9,   c(1, 5, 2))
y <- k_reshape(10:19, c(1, 2, 5))
layer_dot(x, y, axes=c(2, 3))
```

Usage in a Keras model:

```{r}
x1 <- k_reshape(0:9, c(5, 2)) |> layer_dense(8)
x2 <- k_reshape(10:19, c(5, 2)) |> layer_dense(8)
x1$shape
x2$shape
y <- layer_dot(x1, x2, axes=2)
y$shape
```

@returns
    A tensor, the dot product of the samples from the inputs.

@param axes Integer or list of integers, axis or axes along which to
    take the dot product. If a list, should be two integers
    corresponding to the desired axis from the first input and the
    desired axis from the second input, respectively. Note that the
    size of the two selected axes must match.
@param normalize Whether to L2-normalize samples along the dot product axis
    before taking the dot product. If set to `TRUE`, then
    the output of the dot product is the cosine proximity
    between the two samples.
@param ... Standard layer keyword arguments.
@param inputs layers to combine

@export
@family merging layers
@seealso
+ <https:/keras.io/api/layers/merging_layers/dot#dot-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/layers/Dot>
