---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
A `LearningRateSchedule` that uses an exponential decay schedule.

@description
When training a model, it is often useful to lower the learning rate as
the training progresses. This schedule applies an exponential decay function
to an optimizer step, given a provided initial learning rate.

The schedule is a 1-arg callable that produces a decayed learning
rate when passed the current optimizer step. This can be useful for changing
the learning rate value across different invocations of optimizer functions.
It is computed as:

```{r}
decayed_learning_rate <- function(step) {
  initial_learning_rate * decay_rate ^ (step / decay_steps)
}
```

If the argument `staircase` is `TRUE`, then `step / decay_steps` is
an integer division and the decayed learning rate follows a
staircase function.

You can pass this schedule directly into a `optimizer`
as the learning rate.

# Examples
When fitting a Keras model, decay every 100000 steps with a base
of 0.96:

```{r, eval=FALSE}
initial_learning_rate <- 0.1
lr_schedule <- learning_rate_schedule_exponential_decay(
    initial_learning_rate,
    decay_steps=100000,
    decay_rate=0.96,
    staircase=TRUE)

model %>% compile(
  optimizer = optimizer_sgd(learning_rate = lr_schedule),
  loss = 'sparse_categorical_crossentropy',
  metrics = c('accuracy'))

model %>% fit(data, labels, epochs=5)
```

The learning rate schedule is also serializable and deserializable using
`keras$optimizers$schedules$serialize` and
`keras$optimizers$schedules$deserialize`.

@returns
A 1-arg callable learning rate schedule that takes the current optimizer
step and outputs the decayed learning rate, a scalar tensor of the
same type as `initial_learning_rate`.

@param initial_learning_rate
A float. The initial learning rate.

@param decay_steps
A integer. Must be positive. See the decay
computation above.

@param decay_rate
A float. The decay rate.

@param staircase
Boolean.  If `TRUE` decay the learning rate at discrete
intervals.

@param name
String.  Optional name of the operation.  Defaults to
`"ExponentialDecay`".

@export
@family learning rate schedule optimizers
@family schedule optimizers
@seealso
+ <https:/keras.io/api/optimizers/learning_rate_schedules/exponential_decay#exponentialdecay-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/optimizers/schedules/ExponentialDecay>

