---
title: "Writing Custom Keras Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Custom Keras Models} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "Custom Models"
    identifier: "keras-custom-models"
    parent: "keras-advanced"
    weight: 60
---

```{r setup, include = FALSE}
library(keras)
knitr::opts_chunk$set(comment = NA, eval = FALSE)
```

In addition to [sequential models](sequential_model.html) and models created with the [functional API](functional_api.html), you may also define models by creating a custom R6 class and writing your own own `call()` forward pass.

The example below illustrates the implementation of a simple custom custom model that implements a multi-layer-perceptron with optional dropout and batch normalization.

## KerasModel R6 Class

To create a custom Keras model, you create an R6 class derived from `KerasModel`, and implement a `call()` method which defines the model's forward-pass behavior:

```{r}
library(keras)

SimpleMLP <- R6::R6Class("SimpleMLP",
                                  
  inherit = KerasModel,
  
  public = list(
    
    num_classes = NULL,
    use_bn = NULL,
    use_dp = NULL,
    dense1 = NULL,
    dense2 = NULL,
    dp = NULL,
    bn = NULL,
    
    initialize = function(num_classes, use_bn = FALSE, use_dp = FALSE) {
      self$num_classes <- num_classes
      self$use_bn <- use_bn
      self$use_dp <- use_dp
      self$dense1 <- layer_dense(units = 32, activation = "relu")
      self$dense2 <- layer_dense(units = num_classes, activation = "softmax")
      if (self$use_dp)
        self$dp <- layer_dropout(rate = 0.5)
      if (self$use_bn)
        self$bn <- layer_batch_normalization(axis = -1)
    },
    
    call = function(inputs, mask = NULL) {
      x <- self$dense1(inputs)
      if (self$use_dp)
        x <- self$dp(x)
      if (self$use_bn)
        x <- self$bn(x)
      self$dense2(x)
    }
  )
)
```

Layers are defined within the `initialize()` method and the forward pass is specified in the `call()` method. In `call()`, you may specify custom losses by calling `self$add_loss()`, like you would in a custom layer (you can also access any other members of the base Keras model you need to using `self$`).


## Model Wrapper Function

In order to use the custom model, you also need to create a wrapper function which instantiates the model using the `keras_model_custom()` function. For example:

```{r}
# define model wrapper function
keras_model_simple_mlp <- function(num_classes, use_bn = FALSE, use_dp = FALSE, name = NULL) {
  keras_model_custom(SimpleMLP, 
    num_classes = num_classes,
    use_bn = use_bn,
    use_dp = use_dp,
    name = name
  )
}

# create the model 
model <- keras_model_simple_mlp(10, use_dp = TRUE)

# compile graph
model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# Generate dummy data
data <- matrix(runif(1000*100), nrow = 1000, ncol = 100)
labels <- matrix(round(runif(1000, min = 0, max = 9)), nrow = 1000, ncol = 1)

# Convert labels to categorical one-hot encoding
one_hot_labels <- to_categorical(labels, num_classes = 10)

# Train the model
model %>% fit(data, one_hot_labels, epochs=10, batch_size=32)
```






