---
title: "Guide to the Functional API"
author: "JJ Allaire"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: |
  %\VignetteIndexEntry{Guide to the Functional API} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(keras)
```


The Keras functional API is the way to go for defining complex models, such as multi-output models, directed acyclic graphs, or models with shared layers.

This guide assumes that you are already familiar with the [Sequential](sequential_model.html) model.

Let's start with something simple.

## First example: a densely-connected network

The [Sequential](sequential_model.html) model is probably a better choice to implement such a network, but it helps to start with something really simple.

To use the functional API, build your input and output layers and then pass them to the `model()` function. This model can be trained just like Keras sequential models.


```{r}
library(keras)

# input layer
inputs <- layer_input(shape = c(784))
 
# outputs compose input + dense layers
predictions <- inputs %>%
  layer_dense(units = 64, activation = 'relu') %>% 
  layer_dense(units = 64, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax')

# create and compile model
model <- model(inputs = inputs, outputs = predictions) %>%
  compile(
    optimizer = 'rmsprop',
    loss = 'categorical_crossentropy',
    metrics = c('accuracy')
  )
```

## Models can be treated like layers

With the functional API, it is easy to re-use trained models: you can treat any model as if it were a layer. Note that you aren't just re-using the architecture of the model, you are also re-using its weights.

This can allow you to for instance quickly create models that can process sequences of inputs. You could turn an image classification model into a video classification model, in just one line:

```{r}

# Input tensor for sequences of 20 timesteps,
# each containing a 784-dimensional vector
input_sequences <- layer_input(shape = c(20, 784))

# This applies our previous model to every timestep in the input sequences.
# the output of the previous model was a 10-way softmax,
# so the output of the layer below will be a sequence of 20 vectors of size 10.
processed_sequences <- input_sequences %>% 
    time_distributed(model) 
```








