---
title: "TensorFlow Hub with Keras"
output: 
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{TensorFlow Hub with Keras} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "TensorFlow Hub with Keras"
    identifier: "hub-with-keras"
    parent: "keras-using-keras"
    weight: 70
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


[TensorFlow Hub](http://tensorflow.org/hub) is a way to share pretrained model components. See the TensorFlow Module Hub for a searchable listing of pre-trained models. This tutorial demonstrates:

1. How to use TensorFlow Hub with Keras.
2. How to do image classification using TensorFlow Hub.
3. How to do simple transfer learning.

## Setup

```{r}
library(keras)
```

## An ImageNet classifier

### Download the classifier

Use `layer_hub` to load a mobilenet and transform it into a Keras layer. 
Any TensorFlow 2 compatible image classifier URL from tfhub.dev will work here.

```{r}
classifier_url <- "https://tfhub.dev/google/tf2-preview/mobilenet_v2/classification/2" 

mobilenet_layer <- layer_hub(classifier_url, input_shape = shape(224, 224, 3))
```






