---
title: "Python Subclasses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Python Subclasses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "Python Subclasses"
    identifier: "python-classes"
    parent: "keras-advanced-top"
    weight: 10
editor_options:
  markdown:
    wrap: sentence
aliases:
  - /keras/articles/custom_python_classes.html
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(reticulate)
library(keras)
```

When using keras, a desire to create Python-based subclasses can arise in a number of ways.
For example, when you want to:

-   define custom layers and/or models
-   implement custom training logic
-   create custom losses or optimizers
-   define custom callbacks
-   ... and more!

In such scenarios, the most powerful and flexible approach is to directly inherit from, and then modify and/or enhance an appropriate Python class.

Subclassing a Python class in R is generally straightforward.
Two syntaxes are provided: one that adheres to R conventions and uses `R6::R6Class` as the class constructor, and one that adheres more to Python conventions, and attempts to replicate Python syntax in R.

## Examples

### A custom constraint (R6)

For demonstration purposes, let's say you want to implement a custom keras kernel constraint via subclassing.
Using R6:

```{r}
NonNegative <- R6::R6Class("NonNegative",
  inherit = keras$constraints$Constraint,
  public = list(
    "__call__" = function(x) {
       w * k_cast(w >= 0, k_floatx())
    }
  )
)
NonNegative <- r_to_py(NonNegative, convert=TRUE)
```

The `r_to_py` method will convert an R6 class generator into a Python class generator.
After conversion, Python class generators will be different from R6 class generators in a few ways:

-   New class instances are generated by calling the class directly: `NonNegative()` (not `NonNegative$new()`)

-   All methods (functions) are (potentially) modified to ensure their first argument is `self`.

-   All methods have in scope `__class__`, `super` and the class name (`NonNegative`).

-   For convenience, some method names are treated as aliases:

    -   `initialize` is treated as an alias for `__init__`()
    -   `finalize` is treated as an alias for `__del__`()

-   `super` can be accessed in 3 ways:

    1)  R6 style, which supports only single inheritance (the most common type)

    ``` {.r}
    super$initialize()
    ```

    2)  Python 2 style, which requires explicitly providing the class generator and instance

    ``` {.r}
    super(NonNegative, self)$`__init__`()
    ```

    3)  Python 3 style

    ``` {.r}
    super()$`__init__`()
    ```

-   When subclassing Keras base classes, it is generally your responsibility to call `super$initialize()` if you are masking a superclass initializer by providing your own `initialize` method.

-   Passing `convert=FALSE` to `r_to_py()` will mean that all R methods will receive Python objects as arguments, and are expected to return Python objects.
    This allows for some features not available with `convert=TRUE`, namely, modifying some Python objects, like dictionaries or lists, in-place.

-   Active bindings (methods supplied to `R6Class(active=...)`) are converted to Python `@property`-decorated methods.

-   R6 classes with private methods or attributes are not supported.

-   The argument supplied to `inherit` can be:

    -   missing or `NULL`
    -   a Python class generator
    -   an R6 ClassGenerator, as long as it can be converted to a Python class generator as well
    -   a list of Python/R6 classes (for multiple inheritance)
    -   A list of superclasses, with optional additional keywords (e.g., `metaclass=`, only for advanced Python use cases)

### A custom constraint (`%py_class%`)

As an alternative to `r_to_py(R6Class(...))`, we also provide `%py_class%`, a more concise alternative syntax for achieving the same outcome.
`%py_class%` is heavily inspired by the Python syntax for the `class` statement in R, and is especially convenient when translating Python code to R.
Translating the above example, you could write the same using `%py_class%`:

```{r}
NonNegative(keras$constraints$Constraint) %py_class% {
  "__call__" = function(x) {
    w * k_cast(w >= 0, k_floatx())
  }
}
```

Notice, this is very similar to the equivalent [Python code](https://www.tensorflow.org/versions/r2.5/api_docs/python/tf/keras/constraints/Constraint):

```{python, eval=FALSE}
class NonNegative(tf.keras.constraints.Constraint):
    def __call__(self, w):
        return w * tf.cast(tf.math.greater_equal(w, 0.), w.dtype)
```

Some (potentially surprising) notes about `%py_class%`:

-   Just like the Python `class` statement, it assigns the constructed class in the current scope!
    (There is no need to write `NonNegative <- ...`).

-   The left hand side can be:

    -   A bare symbol, `ClassName`
    -   A pseudo-call, with superclasses and keywords as arguments: `ClassName(Superclass1, Superclass2, metaclass=my_metaclass)`

-   The right hand side is evaluated in a new environment to form the namespace for the class methods.

-   One keyword treated specially is `convert`.
    If you want to call `r_to_py` with `convert=FALSE`, pass it as a keyword:

```{r, eval=FALSE}
NonNegative(keras$constraints$Constraint, convert=FALSE) %py_class% { ... }
```

-   An additional convenience is that if the first expression of a function body
    or the class body is a literal character string, it is automatically taken as
    the `__doc__` attribute of the class or method. The doc string will then be
    visible to both python and R tools e.g. `reticulate::py_help()`. 
    See `?py_class` for an example.

In all other regards, `%py_class%` is equivalent to `r_to_py(R6Class())` (indeed, under the hood, they do the same thing).

### A custom layer (R6)

The same pattern can be extended to all sorts of keras objects.
For example, a custom layer can be written by subclassing the base Keras Layer:

```{r}
CustomLayer <- R6::R6Class("CustomLayer",

  inherit = keras$layers$Layer,

  public = list(

    initialize = function(output_dim) {
      self$output_dim <- output_dim
    },

    build = function(input_shape) {
      self$kernel <- self$add_weight(
        name = 'kernel',
        shape = list(input_shape[[2]], self$output_dim),
        initializer = initializer_random_normal(),
        trainable = TRUE
      )
    },

    call = function(x, mask = NULL) {
      k_dot(x, self$kernel)
    },

    compute_output_shape = function(input_shape) {
      list(input_shape[[1]], self$output_dim)
    }
  )
)
```

### A custom layer (`%py_class%`)

or using `%py_class%`:

```{r}
CustomLayer(keras$layers$Layer) %py_class% {

  initialize = function(output_dim) {
    self$output_dim <- output_dim
  }

  build = function(input_shape) {
    self$kernel <- self$add_weight(
      name = 'kernel',
      shape = list(input_shape[[2]], self$output_dim),
      initializer = initializer_random_normal(),
      trainable = TRUE
    )
  }

  call = function(x, mask = NULL) {
    k_dot(x, self$kernel)
  }

  compute_output_shape = function(input_shape) {
    list(input_shape[[1]], self$output_dim)
  }
}
```
