---
title: "Tutorial: Save and Restore Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Save and Restore Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "Save and Restore Models"
    identifier: "keras-tutorial-save-restore"
    parent: "keras-tutorials"
    weight: 50
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


Model progress can be saved after as well as during training. This means a model can resume where it left off and avoid long training times. Saving also means you can share your model and others can recreate your work. When publishing research models and techniques, most machine learning practitioners share:

- code to create the model, and
- the trained weights, or parameters, for the model

Sharing this data helps others understand how the model works and try it themselves with new data.

## Setup

We'll use the [MNIST](http://yann.lecun.com/exdb/mnist/) dataset to train our model to demonstrate saving weights. To speed up these demonstration runs, only use the first 1000 examples:


```{r}
library(keras)

mnist <- dataset_mnist()

c(train_images, train_labels) %<-% mnist$train
c(test_images, test_labels) %<-% mnist$test

train_labels <- train_labels[1:1000]
test_labels <- test_labels[1:1000]

train_images <- train_images[1:1000, , ] %>%
  array_reshape(c(1000, 28 * 28)) %>%
  `/`(., 255)
test_images <- test_images[1:1000, , ] %>%
  array_reshape(c(1000, 28 * 28)) %>%
  `/`(., 255)

```



## Define a model

Let's build a simple model we'll use to demonstrate saving and loading weights.


```{r}
# Returns a short sequential model
create_model <- function() {
  model <- keras_model_sequential() %>%
    layer_dense(units = 512, activation = "relu", input_shape = 784) %>%
    layer_dropout(0.2) %>%
    layer_dense(units = 10, activation = "softmax")
  model %>% compile(
    optimizer = "adam",
    loss = "sparse_categorical_crossentropy",
    metrics = list("accuracy")
  )
  model
}

model <- create_model()
model %>% summary()

```

```
__________________________________________________________________________________________________________
Layer (type)                                   Output Shape                               Param #         
==========================================================================================================
dense_1 (Dense)                                (None, 512)                                401920          
__________________________________________________________________________________________________________
dropout_1 (Dropout)                            (None, 512)                                0               
__________________________________________________________________________________________________________
dense_2 (Dense)                                (None, 10)                                 5130            
==========================================================================================================
Total params: 407,050
Trainable params: 407,050
Non-trainable params: 0
```

## Save the entire model

The habitual form of saving a Keras model is saving to the HDF5 format.

The resulting file contains the weight values, the model's configuration, and even the optimizer's configuration. This allows you to save a model and resume training later — from the exact same state — without access to the original code.

```{r}
model <- create_model()

model %>% fit(train_images, train_labels, epochs = 5)

model %>% save_model_hdf5("my_model.h5")
```


If you only wanted to save the weights, you could replace that last line by 

```{r}
model %>% save_model_weights_hdf5("my_model_weights.h5")
```

Now recreate the model from that file:

```{r}
new_model <- load_model_hdf5("my_model.h5")
new_model %>% summary()
```

```
__________________________________________________________________________________________________________
Layer (type)                                   Output Shape                               Param #         
==========================================================================================================
dense_1 (Dense)                                (None, 512)                                401920          
__________________________________________________________________________________________________________
dropout_1 (Dropout)                            (None, 512)                                0               
__________________________________________________________________________________________________________
dense_2 (Dense)                                (None, 10)                                 5130            
==========================================================================================================
Total params: 407,050
Trainable params: 407,050
Non-trainable params: 0
__________________________________________________________________________________________________________
```



## Save checkpoints during training

It is useful to automatically save checkpoints during and at the end of training. This way you can use a trained model without having to retrain it, or pick-up training where you left of, in case the training process was interrupted.

`callback_model_checkpoint` is a callback that performs this task. 

The callback takes a couple of arguments to configure checkpointing. By default, `save_weights_only` is set to false, which means the complete model is being saved - including architecture and configuration. You can then restore the model as outlined in the previous paragraph.

Now here, let's focus on just saving and restoring weights. In the following code snippet, we are setting `save_weights_only` to true, so we will need the model definition on restore.

The `filepath` argument can contain named formatting options, for example: if `filepath` is `weights.{epoch:02d}-{val_loss:.2f}.hdf5`, then the model checkpoints will be saved with the epoch number and the validation loss in the filename. 

The saved model weights again will be in HDF5 format.


## Checkpoint callback usage

Train the model and pass it the `callback_model_checkpoint`:

```{r}
checkpoint_dir <- "checkpoints"
dir.create(checkpoint_dir, showWarnings = FALSE)
filepath <- file.path(checkpoint_dir, "weights.{epoch:02d}-{val_loss:.2f}.hdf5")

# Create checkpoint callback
cp_callback <- callback_model_checkpoint(
  filepath = filepath,
  save_weights_only = TRUE,
  verbose = 1
)

model <- create_model()

model %>% fit(
  train_images,
  train_labels,
  epochs = 10, 
  validation_data = list(test_images, test_labels),
  callbacks = list(cp_callback)  # pass callback to training
)
```


Inspect the files that were created:

```{r}
list.files(checkpoint_dir)
```

```
 [1] "weights.01-0.67.hdf5" "weights.02-0.57.hdf5" "weights.03-0.47.hdf5" "weights.04-0.44.hdf5"
 [5] "weights.05-0.43.hdf5" "weights.06-0.39.hdf5" "weights.07-0.40.hdf5" "weights.08-0.39.hdf5"
 [9] "weights.09-0.44.hdf5" "weights.10-0.39.hdf5"
 ```
 


Create a new, untrained model. When restoring a model from only weights, you must have a model with the same architecture as the original model. Since it's the same model architecture, we can share weights despite that it's a different instance of the model.

Now rebuild a fresh, untrained model, and evaluate it on the test set. An untrained model will perform at chance levels (~10% accuracy):

```{r}
fresh_model <- create_model()
c(test_loss, test_acc) %<-% (fresh_model %>% evaluate(test_images, test_labels))

paste('Test loss:', test_loss)
paste('Test accuracy:', test_acc)
```

```
1000/1000 [==============================] - 0s 255us/step
[1] "Test loss: 2.34574563026428"
[1] "Test accuracy: 0.096"
```

Then load the weights from the latest checkpoint (epoch 10), and re-evaluate:

```{r}
fresh_model %>% load_model_weights_hdf5(file.path(checkpoint_dir, "weights.10-0.39.hdf5"))
c(test_loss, test_acc) %<-% (fresh_model %>% evaluate(test_images, test_labels))

paste('Test loss:', test_loss)
paste('Test accuracy:', test_acc)
```

```
1000/1000 [==============================] - 0s 34us/step
[1] "Test loss: 0.394947263240814"
[1] "Test accuracy: 0.873"
```

To reduce the number of files, you can also save model weights only once every $n$th epoch. E.g.,

```{r}
checkpoint_dir <- "checkpoints"
dir.create(checkpoint_dir, showWarnings = FALSE)
filepath <- file.path(checkpoint_dir, "weights.{epoch:02d}-{val_loss:.2f}.hdf5")

# Create checkpoint callback
cp_callback <- callback_model_checkpoint(
  filepath = filepath,
  save_weights_only = TRUE,
  period = 5,
  verbose = 1
)

model <- create_model()

model %>% fit(
  train_images,
  train_labels,
  epochs = 10, 
  validation_data = list(test_images, test_labels),
  callbacks = list(cp_callback)  # pass callback to training
)

list.files(checkpoint_dir)
```

```
[1] "weights.05-0.43.hdf5" "weights.10-0.41.hdf5"
```

Alternatively, you can also decide to save only the best model, where best by default is defined as validation loss.
See the [documentation for callback_model_checkpoint](https://tensorflow.rstudio.com/keras/reference/callback_model_checkpoint.html) for further information.

```{r}
checkpoint_dir <- "checkpoints"
dir.create(checkpoint_dir, showWarnings = FALSE)
filepath <- file.path(checkpoint_dir, "weights.{epoch:02d}-{val_loss:.2f}.hdf5")

# Create checkpoint callback
cp_callback <- callback_model_checkpoint(
  filepath = filepath,
  save_weights_only = TRUE,
  save_best_only = TRUE,
  verbose = 1
)

model <- create_model()

model %>% fit(
  train_images,
  train_labels,
  epochs = 10, 
  validation_data = list(test_images, test_labels),
  callbacks = list(cp_callback)  # pass callback to training
)

list.files(checkpoint_dir)
```

```
[1] "weights.01-0.71.hdf5" "weights.02-0.54.hdf5" "weights.03-0.46.hdf5" "weights.04-0.45.hdf5"
[5] "weights.05-0.42.hdf5" "weights.08-0.42.hdf5" "weights.09-0.41.hdf5" "weights.10-0.41.hdf5"
```

In this case, weights were saved on all epochs but the 6th and 7th, where validation loss did not improve.
