---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the squared hinge loss between `y_true` & `y_pred`.

@description
Formula:

```{r, eval=FALSE}
loss <- square(maximum(1 - y_true * y_pred, 0))
```

`y_true` values are expected to be -1 or 1. If binary (0 or 1) labels are
provided we will convert them to -1 or 1.

# Examples
```{r}
y_true <- array(sample(c(-1,1), 6, replace = TRUE), dim = c(2, 3))
y_pred <- random_uniform(c(2, 3))
loss <- loss_squared_hinge(y_true, y_pred)
```

@returns
Squared hinge loss values with shape = `[batch_size, d0, .. dN-1]`.

@param reduction
Type of reduction to apply to the loss. In almost all cases
this should be `"sum_over_batch_size"`.
Supported options are `"sum"`, `"sum_over_batch_size"` or `NULL`.

@param name
Optional name for the loss instance.

@param y_true
The ground truth values. `y_true` values are expected to be -1
or 1. If binary (0 or 1) labels are provided we will convert them
to -1 or 1 with shape = `[batch_size, d0, .. dN]`.

@param y_pred
The predicted values with shape = `[batch_size, d0, .. dN]`.

@param ...
Passed on to the Python callable

@export
@family losses
@seealso
+ <https:/keras.io/api/losses/hinge_losses#squaredhinge-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/losses/SquaredHinge>

