---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Approximates the AUC (Area under the curve) of the ROC or PR curves.

@description
The AUC (Area under the curve) of the ROC (Receiver operating
characteristic; default) or PR (Precision Recall) curves are quality
measures of binary classifiers. Unlike the accuracy, and like cross-entropy
losses, ROC-AUC and PR-AUC evaluate all the operational points of a model.

This class approximates AUCs using a Riemann sum. During the metric
accumulation phrase, predictions are accumulated within predefined buckets
by value. The AUC is then computed by interpolating per-bucket averages.
These buckets define the evaluated operational points.

This metric creates four local variables, `true_positives`,
`true_negatives`, `false_positives` and `false_negatives` that are used to
compute the AUC.  To discretize the AUC curve, a linearly spaced set of
thresholds is used to compute pairs of recall and precision values. The area
under the ROC-curve is therefore computed using the height of the recall
values by the false positive rate, while the area under the PR-curve is the
computed using the height of the precision values by the recall.

This value is ultimately returned as `auc`, an idempotent operation that
computes the area under a discretized curve of precision versus recall
values (computed using the aforementioned variables). The `num_thresholds`
variable controls the degree of discretization with larger numbers of
thresholds more closely approximating the true AUC. The quality of the
approximation may vary dramatically depending on `num_thresholds`. The
`thresholds` parameter can be used to manually specify thresholds which
split the predictions more evenly.

For a best approximation of the real AUC, `predictions` should be
distributed approximately uniformly in the range `[0, 1]` (if
`from_logits=FALSE`). The quality of the AUC approximation may be poor if
this is not the case. Setting `summation_method` to 'minoring' or 'majoring'
can help quantify the error in the approximation by providing lower or upper
bound estimate of the AUC.

If `sample_weight` is `NULL`, weights default to 1.
Use `sample_weight` of 0 to mask values.

# Usage
Standalone usage:

```{r}
m <- metric_auc(num_thresholds = 3)
m$update_state(c(0, 0, 1, 1), c(0, 0.5, 0.3, 0.9))
# threshold values are [0 - 1e-7, 0.5, 1 + 1e-7]
# tp = [2, 1, 0], fp = [2, 0, 0], fn = [0, 1, 2], tn = [0, 2, 2]
# tp_rate = recall = [1, 0.5, 0], fp_rate = [1, 0, 0]
# auc = ((((1 + 0.5) / 2) * (1 - 0)) + (((0.5 + 0) / 2) * (0 - 0)))
#     = 0.75
m$result()
# 0.75
```

```{r}
m$reset_state()
m$update_state(c(0, 0, 1, 1), c(0, 0.5, 0.3, 0.9),
               sample_weight=c(1, 0, 0, 1))
m$result()
# 1.0
```

Usage with `compile()` API:

```{r, eval = FALSE}
# Reports the AUC of a model outputting a probability.
model %>% compile(optimizer = 'sgd',
                  loss = loss_binary_crossentropy(),
                  metrics = list(metric_auc()))

# Reports the AUC of a model outputting a logit.
model %>% compile(optimizer = 'sgd',
                  loss = loss_binary_crossentropy(from_logits = TRUE),
                  metrics = list(metric_auc(from_logits = TRUE)))
```

@param num_thresholds (Optional) The number of thresholds to
    use when discretizing the roc curve. Values must be > 1.
    Defaults to `200`.
@param curve (Optional) Specifies the name of the curve to be computed,
    `'ROC'` (default) or `'PR'` for the Precision-Recall-curve.
@param summation_method (Optional) Specifies the [Riemann summation method](
      https://en.wikipedia.org/wiki/Riemann_sum) used.
      'interpolation' (default) applies mid-point summation scheme for
      `ROC`.  For PR-AUC, interpolates (true/false) positives but not
      the ratio that is precision (see Davis & Goadrich 2006 for
      details); 'minoring' applies left summation for increasing
      intervals and right summation for decreasing intervals; 'majoring'
      does the opposite.
@param name (Optional) string name of the metric instance.
@param dtype (Optional) data type of the metric result.
@param thresholds (Optional) A list of floating point values to use as the
    thresholds for discretizing the curve. If set, the `num_thresholds`
    parameter is ignored. Values should be in `[0, 1]`. Endpoint
    thresholds equal to {`-epsilon`, `1+epsilon`} for a small positive
    epsilon value will be automatically included with these to correctly
    handle predictions equal to exactly 0 or 1.
@param multi_label boolean indicating whether multilabel data should be
    treated as such, wherein AUC is computed separately for each label
    and then averaged across labels, or (when `FALSE`) if the data
    should be flattened into a single label before AUC computation. In
    the latter case, when multilabel data is passed to AUC, each
    label-prediction pair is treated as an individual data point. Should
    be set to FALSE for multi-class data.
@param num_labels (Optional) The number of labels, used when `multi_label` is
    TRUE. If `num_labels` is not specified, then state variables get
    created on the first call to `update_state`.
@param label_weights (Optional) list, array, or tensor of non-negative weights
    used to compute AUCs for multilabel data. When `multi_label` is
    TRUE, the weights are applied to the individual label AUCs when they
    are averaged to produce the multi-label AUC. When it's FALSE, they
    are used to weight the individual label predictions in computing the
    confusion matrix on the flattened data. Note that this is unlike
    `class_weights` in that `class_weights` weights the example
    depending on the value of its label, whereas `label_weights` depends
    only on the index of that label before flattening; therefore
    `label_weights` should not be used for multi-class data.
@param from_logits boolean indicating whether the predictions (`y_pred` in
`update_state`) are probabilities or sigmoid logits. As a rule of thumb,
when using a keras loss, the `from_logits` constructor argument of the
loss should match the AUC `from_logits` constructor argument.
@param ... Passed on to the Python callable

@export
@family metric
@seealso
+ <https:/keras.io/api/metrics/classification_metrics#auc-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/metrics/AUC>

