---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the crossentropy metric between the labels and predictions.

@description
This is the crossentropy metric class to be used when there are only two
label classes (0 and 1).

# Examples
Standalone usage:

```{r}
m <- metric_binary_crossentropy()
m$update_state(rbind(c(0, 1), c(0, 0)), rbind(c(0.6, 0.4), c(0.4, 0.6)))
m$result()
```

```{r}
m$reset_state()
m$update_state(rbind(c(0, 1), c(0, 0)), rbind(c(0.6, 0.4), c(0.4, 0.6)),
               sample_weight=c(1, 0))
m$result()
```

Usage with `compile()` API:

```{r, eval = FALSE}
model %>% compile(
    optimizer = 'sgd',
    loss = 'mse',
    metrics = list(metric_binary_crossentropy()))
```

@param name
(Optional) string name of the metric instance.

@param dtype
(Optional) data type of the metric result.

@param from_logits
(Optional) Whether output is expected
to be a logits tensor. By default, we consider
that output encodes a probability distribution.

@param label_smoothing
(Optional) Float in `[0, 1]`.
When > 0, label values are smoothed,
meaning the confidence on label values are relaxed.
e.g. `label_smoothing=0.2` means that we will use
a value of 0.1 for label "0" and 0.9 for label "1".

@param y_true
Ground truth values. shape = `[batch_size, d0, .. dN]`.

@param y_pred
The predicted values. shape = `[batch_size, d0, .. dN]`.

@param axis
The axis along which the mean is computed. Defaults to `-1`.

@param ...
Passed on to the Python callable

@export
@family probabilistic metrics
@family metrics
@seealso
+ <https:/keras.io/api/metrics/probabilistic_metrics#binarycrossentropy-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/metrics/BinaryCrossentropy>

