---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
A preprocessing layer that maps strings to (possibly encoded) indices.

@description
This layer translates a set of arbitrary strings into integer output via a
table-based vocabulary lookup. This layer will perform no splitting or
transformation of input strings. For a layer than can split and tokenize
natural language, see the `layer_text_vectorization` layer.

The vocabulary for the layer must be either supplied on construction or
learned via `adapt()`. During `adapt()`, the layer will analyze a data set,
determine the frequency of individual strings tokens, and create a
vocabulary from them. If the vocabulary is capped in size, the most frequent
tokens will be used to create the vocabulary and all others will be treated
as out-of-vocabulary (OOV).

There are two possible output modes for the layer.
When `output_mode` is `"int"`,
input strings are converted to their index in the vocabulary (an integer).
When `output_mode` is `"multi_hot"`, `"count"`, or `"tf_idf"`, input strings
are encoded into an array where each dimension corresponds to an element in
the vocabulary.

The vocabulary can optionally contain a mask token as well as an OOV token
(which can optionally occupy multiple indices in the vocabulary, as set
by `num_oov_indices`).
The position of these tokens in the vocabulary is fixed. When `output_mode`
is `"int"`, the vocabulary will begin with the mask token (if set), followed
by OOV indices, followed by the rest of the vocabulary. When `output_mode`
is `"multi_hot"`, `"count"`, or `"tf_idf"` the vocabulary will begin with
OOV indices and instances of the mask token will be dropped.

**Note:** This layer uses TensorFlow internally. It cannot
be used as part of the compiled computation graph of a model with
any backend other than TensorFlow.
It can however be used with any backend when running eagerly.
It can also always be used as part of an input preprocessing pipeline
with any backend (outside the model itself), which is how we recommend
to use this layer.

**Note:** This layer is safe to use inside a `tf.data` pipeline
(independently of which backend you're using).

# Examples
**Creating a lookup layer with a known vocabulary**

This example creates a lookup layer with a pre-existing vocabulary.

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c("a", "c", "d"), c("d", "z", "b"))
layer <- layer_string_lookup(vocabulary=vocab)
layer(data)
```

**Creating a lookup layer with an adapted vocabulary**

This example creates a lookup layer and generates the vocabulary by
analyzing the dataset.

```{r}
data <- rbind(c("a", "c", "d"), c("d", "z", "b"))
layer <- layer_string_lookup()
layer %>% adapt(data)
get_vocabulary(layer)
```

Note that the OOV token `"[UNK]"` has been added to the vocabulary.
The remaining tokens are sorted by frequency
(`"d"`, which has 2 occurrences, is first) then by inverse sort order.

```{r}
data <- rbind(c("a", "c", "d"), c("d", "z", "b"))
layer <- layer_string_lookup()
layer %>% adapt(data)
layer(data)
```

**Lookups with multiple OOV indices**

This example demonstrates how to use a lookup layer with multiple OOV
indices.  When a layer is created with more than one OOV index, any OOV
values are hashed into the number of OOV buckets, distributing OOV values in
a deterministic fashion across the set.

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c("a", "c", "d"), c("m", "z", "b"))
layer <- layer_string_lookup(vocabulary = vocab, num_oov_indices = 2)
layer(data)
```

Note that the output for OOV value 'm' is 0, while the output for OOV value
`"z"` is 1. The in-vocab terms have their output index increased by 1 from
earlier examples (a maps to 2, etc) in order to make space for the extra OOV
value.

**One-hot output**

Configure the layer with `output_mode='one_hot'`. Note that the first
`num_oov_indices` dimensions in the ont_hot encoding represent OOV values.

```{r}
vocab <- c("a", "b", "c", "d")
data <- c("a", "b", "c", "d", "z")
layer <- layer_string_lookup(vocabulary = vocab, output_mode = 'one_hot')
layer(data)
```

**Multi-hot output**

Configure the layer with `output_mode='multi_hot'`. Note that the first
`num_oov_indices` dimensions in the multi_hot encoding represent OOV values.

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c("a", "c", "d", "d"), c("d", "z", "b", "z"))
layer <- layer_string_lookup(vocabulary = vocab, output_mode = 'multi_hot')
layer(data)
```

**Token count output**

Configure the layer with `output_mode='count'`. As with multi_hot output,
the first `num_oov_indices` dimensions in the output represent OOV values.

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c("a", "c", "d", "d"), c("d", "z", "b", "z"))
layer <- layer_string_lookup(vocabulary = vocab, output_mode = 'count')
layer(data)
```

**TF-IDF output**

Configure the layer with `output_mode="tf_idf"`. As with multi_hot output,
the first `num_oov_indices` dimensions in the output represent OOV values.

Each token bin will output `token_count * idf_weight`, where the idf weights
are the inverse document frequency weights per token. These should be
provided along with the vocabulary. Note that the `idf_weight` for OOV
values will default to the average of all idf weights passed in.

```{r}
vocab <- c("a", "b", "c", "d")
idf_weights <- c(0.25, 0.75, 0.6, 0.4)
data <- rbind(c("a", "c", "d", "d"), c("d", "z", "b", "z"))
layer <- layer_string_lookup(output_mode = "tf_idf")
layer %>% set_vocabulary(vocab, idf_weights=idf_weights)
layer(data)
```

To specify the idf weights for oov values, you will need to pass the entire
vocabularly including the leading oov token.

```{r}
vocab <- c("[UNK]", "a", "b", "c", "d")
idf_weights <- c(0.9, 0.25, 0.75, 0.6, 0.4)
data <- rbind(c("a", "c", "d", "d"), c("d", "z", "b", "z"))
layer <- layer_string_lookup(output_mode = "tf_idf")
layer %>% set_vocabulary(vocab, idf_weights=idf_weights)
layer(data)
```

When adapting the layer in `"tf_idf"` mode, each input sample will be
considered a document, and IDF weight per token will be calculated as
`log(1 + num_documents / (1 + token_document_count))`.

**Inverse lookup**

This example demonstrates how to map indices to strings using this layer.
(You can also use `adapt()` with `inverse=TRUE`, but for simplicity we'll
pass the vocab in this example.)

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c(1, 3, 4), c(4, 0, 2))
layer <- layer_string_lookup(vocabulary = vocab, invert = TRUE)
layer(data)
```

Note that the first index correspond to the oov token by default.

**Forward and inverse lookup pairs**

This example demonstrates how to use the vocabulary of a standard lookup
layer to create an inverse lookup layer.

```{r}
vocab <- c("a", "b", "c", "d")
data <- rbind(c("a", "c", "d"), c("d", "z", "b"))
layer <- layer_string_lookup(vocabulary = vocab)
i_layer <- layer_string_lookup(vocabulary = vocab, invert = TRUE)
int_data <- layer(data)
i_layer(int_data)
```

In this example, the input value `"z"` resulted in an output of `"[UNK]"`,
since 1000 was not in the vocabulary - it got represented as an OOV, and all
OOV values are returned as `"[UNK]"` in the inverse layer. Also, note that
for the inverse to work, you must have already set the forward layer
vocabulary either directly or via `adapt()` before calling
`get_vocabulary()`.

@param max_tokens Maximum size of the vocabulary for this layer. This should
    only be specified when adapting the vocabulary or when setting
    `pad_to_max_tokens=TRUE`. If NULL, there is no cap on the size of
    the vocabulary. Note that this size includes the OOV
    and mask tokens. Defaults to `NULL`.
@param num_oov_indices The number of out-of-vocabulary tokens to use.
    If this value is more than 1, OOV inputs are modulated to
    determine their OOV value.
    If this value is 0, OOV inputs will cause an error when calling
    the layer. Defaults to `1`.
@param mask_token A token that represents masked inputs. When `output_mode` is
    `"int"`, the token is included in vocabulary and mapped to index 0.
    In other output modes, the token will not appear
    in the vocabulary and instances of the mask token
    in the input will be dropped. If set to `NULL`,
    no mask term will be added. Defaults to `NULL`.
@param oov_token Only used when `invert` is TRUE. The token to return for OOV
    indices. Defaults to `"[UNK]"`.
@param vocabulary Optional. Either an array of integers or a string path to a
    text file. If passing an array, can pass a list, list,
    1D NumPy array, or 1D tensor containing the integer vocbulary terms.
    If passing a file path, the file should contain one line per term
    in the vocabulary. If this argument is set,
    there is no need to `adapt()` the layer.
@param vocabulary_dtype The dtype of the vocabulary terms, for example
    `"int64"` or `"int32"`. Defaults to `"int64"`.
@param idf_weights Only valid when `output_mode` is `"tf_idf"`.
    A list, list, 1D NumPy array, or 1D tensor or the same length
    as the vocabulary, containing the floating point inverse document
    frequency weights, which will be multiplied by per sample term
    counts for the final TF-IDF weight.
    If the `vocabulary` argument is set, and `output_mode` is
    `"tf_idf"`, this argument must be supplied.
@param invert Only valid when `output_mode` is `"int"`.
    If `TRUE`, this layer will map indices to vocabulary items
    instead of mapping vocabulary items to indices.
    Defaults to `FALSE`.
@param output_mode Specification for the output of the layer. Values can be
    `"int"`, `"one_hot"`, `"multi_hot"`, `"count"`, or `"tf_idf"`
    configuring the layer as follows:
    - `"int"`: Return the vocabulary indices of the input tokens.
    - `"one_hot"`: Encodes each individual element in the input into an
        array the same size as the vocabulary,
        containing a 1 at the element index. If the last dimension
        is size 1, will encode on that dimension.
        If the last dimension is not size 1, will append a new
        dimension for the encoded output.
    - `"multi_hot"`: Encodes each sample in the input into a single
        array the same size as the vocabulary,
        containing a 1 for each vocabulary term present in the sample.
        Treats the last dimension as the sample dimension,
        if input shape is `(..., sample_length)`,
        output shape will be `(..., num_tokens)`.
    - `"count"`: As `"multi_hot"`, but the int array contains
        a count of the number of times the token at that index
        appeared in the sample.
    - `"tf_idf"`: As `"multi_hot"`, but the TF-IDF algorithm is
        applied to find the value in each token slot.
    For `"int"` output, any shape of input and output is supported.
    For all other output modes, currently only output up to rank 2
    is supported. Defaults to `"int"`.
@param pad_to_max_tokens Only applicable when `output_mode` is `"multi_hot"`,
    `"count"`, or `"tf_idf"`. If `TRUE`, the output will have
    its feature axis padded to `max_tokens` even if the number
    of unique tokens in the vocabulary is less than `max_tokens`,
    resulting in a tensor of shape `(batch_size, max_tokens)`
    regardless of vocabulary size. Defaults to `FALSE`.
@param sparse Boolean. Only applicable to `"multi_hot"`, `"count"`, and
    `"tf_idf"` output modes. Only supported with TensorFlow
    backend. If `TRUE`, returns a `SparseTensor`
    instead of a dense `Tensor`. Defaults to `FALSE`.
@param encoding Optional. The text encoding to use to interpret the input
    strings. Defaults to `"utf-8"`.
@param object Object to compose the layer with. A tensor, array, or sequential model.
@param name String, name for the object
@param ... Passed on to the Python callable

@export
@family preprocessing layers
@seealso
+ <https:/keras.io/api/layers/preprocessing_layers/categorical/string_lookup#stringlookup-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/layers/StringLookup>

