---
title: 'Imbalanced classification: credit card fraud detection'
author: '[fchollet](https://twitter.com/fchollet)'
date-created: 2019/05/28
last-modified: 2020/04/17
description: Demonstration of how to handle highly imbalanced classification problems.
accelerator: GPU
output: rmarkdown::html_vignette
domain: structured data
category: basic
backend: jax
knit: ({source(here::here("tools/knit.R")); knit_vignette})
---

```{r setup}
library(keras3)
use_backend("jax")
```
## Introduction

This example looks at the
[Kaggle Credit Card Fraud Detection](https://www.kaggle.com/mlg-ulb/creditcardfraud/)
dataset to demonstrate how
to train a classification model on data with highly imbalanced classes.
You can download the data by clicking "Download" at
the link, or if you're setup with a kaggle API key at
`"~/.kaggle/kagle.json"`, you can run the following:

```{r, eval = FALSE}
reticulate::py_install("kaggle", pip = TRUE)
system("kaggle datasets download -d mlg-ulb/creditcardfraud")
zip::unzip("creditcardfraud.zip", files = "creditcard.csv")
```

## First, load the data

```{r, message=FALSE}
library(readr)
df <- read_csv("creditcard.csv", col_types = cols(
  Class = col_integer(),
  .default = col_double()
))
tibble::glimpse(df)
```

## Prepare a validation set

```{r}
val_idx <- nrow(df) %>% sample.int(., round( . * 0.2))
val_df <- df[val_idx, ]
train_df <- df[-val_idx, ]

cat("Number of training samples:", nrow(train_df), "\n")
cat("Number of validation samples:", nrow(val_df), "\n")
```

## Analyze class imbalance in the targets

```{r}
counts <- table(train_df$Class)
counts

cat(sprintf("Number of positive samples in training data: %i (%.2f%% of total)",
            counts["1"], 100 * counts["1"] / sum(counts)))

weight_for_0 = 1 / counts["0"]
weight_for_1 = 1 / counts["1"]
```

## Normalize the data using training set statistics

```{r}
feature_names <- colnames(train_df) %>% setdiff("Class")

train_features <- as.matrix(train_df[feature_names])
train_targets <- as.matrix(train_df$Class)

val_features <- as.matrix(val_df[feature_names])
val_targets <- as.matrix(val_df$Class)

train_features %<>% scale()
val_features %<>% scale(center = attr(train_features, "scaled:center"),
                        scale = attr(train_features, "scaled:scale"))
```

## Build a binary classification model

```{r}
model <-
  keras_model_sequential(input_shape = ncol(train_features)) |>
  layer_dense(256, activation = "relu") |>
  layer_dense(256, activation = "relu") |>
  layer_dropout(0.3) |>
  layer_dense(256, activation = "relu") |>
  layer_dropout(0.3) |>
  layer_dense(1, activation = "sigmoid")

model
```

## Train the model with `class_weight` argument

```{r}
metrics <- list(
  metric_false_negatives(name = "fn"),
  metric_false_positives(name = "fp"),
  metric_true_negatives(name = "tn"),
  metric_true_positives(name = "tp"),
  metric_precision(name = "precision"),
  metric_recall(name = "recall")
)
model |> compile(
  optimizer = optimizer_adam(1e-2),
  loss = "binary_crossentropy",
  metrics = metrics
)
callbacks <- list(
  callback_model_checkpoint("fraud_model_at_epoch_{epoch}.keras")
)

class_weight <- list("0" = weight_for_0,
                     "1" = weight_for_1)

model |> fit(
  train_features, train_targets,
  validation_data = list(val_features, val_targets),
  class_weight = class_weight,
  batch_size = 2048,
  epochs = 30,
  callbacks = callbacks,
  verbose = 2
)
```

```{r}
val_pred <- model %>%
  predict(val_features) %>%
  { as.integer(. > 0.5) }

pred_correct <- val_df$Class == val_pred
cat(sprintf("Validation accuracy: %.2f", mean(pred_correct)))

fraudulent <- val_df$Class == 1

n_fraudulent_detected <- sum(fraudulent & pred_correct)
n_fraudulent_missed <- sum(fraudulent & !pred_correct)
n_legitimate_flagged <- sum(!fraudulent & !pred_correct)
```

## Conclusions

At the end of training, out of
`r prettyNum(nrow(val_df), big.mark = ",")` validation transactions, we
are:

- Correctly identifying
  `r prettyNum(n_fraudulent_detected, big.mark = ",")` of them as
  fraudulent
- Missing `r prettyNum(n_fraudulent_missed, big.mark = ",")`
  fraudulent transactions
- At the cost of incorrectly flagging
  `r prettyNum(n_legitimate_flagged, big.mark = ",")` legitimate
  transactions

In the real world, one would put an even higher weight on class 1,
so as to reflect that False Negatives are more costly than False Positives.

Next time your credit card gets declined in an online purchase -- this is why.

<!-- | Trained Model                                                                                                                                                          | Demo                                                                                                                                                                             | -->
<!-- |------------------------------------------------------------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------| -->
<!-- | [![Generic badge](https://img.shields.io/badge/%F0%9F%A4%97%20Model-Imbalanced%20Classification-black.svg)](https://huggingface.co/keras-io/imbalanced_classification) | [![Generic badge](https://img.shields.io/badge/%F0%9F%A4%97%20Spaces-Imbalanced%20Classification-black.svg)](https://huggingface.co/spaces/keras-io/Credit_Card_Fraud_Detection) | -->

<!-- : Example available on HuggingFace. -->
