---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the crossentropy metric between the labels and predictions.

@description
This is the crossentropy metric class to be used when there are multiple
label classes (2 or more). It assumes that labels are one-hot encoded,
e.g., when labels values are `c(2, 0, 1)`, then
`y_true` is `rbind(c([0, 0, 1), c(1, 0, 0), c(0, 1, 0))`.

# Examples
Standalone usage:

```{r}
# EPSILON = 1e-7, y = y_true, y` = y_pred
# y` = clip_k_clip_by_value(output, EPSILON, 1. - EPSILON)
# y` = rbind(c(0.05, 0.95, EPSILON), c(0.1, 0.8, 0.1))
# xent = -sum(y * log(y'), axis = -1)
#      = -((log 0.95), (log 0.1))
#      = [0.051, 2.302]
# Reduced xent = (0.051 + 2.302) / 2
m <- metric_categorical_crossentropy()
m$update_state(rbind(c(0, 1, 0), c(0, 0, 1)),
               rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1)))
m$result()
# 1.1769392
```

```{r}
m$reset_state()
m$update_state(rbind(c(0, 1, 0), c(0, 0, 1)),
               rbind(c(0.05, 0.95, 0), c(0.1, 0.8, 0.1)),
               sample_weight = c(0.3, 0.7))
m$result()
```

Usage with `compile()` API:

```{r, eval = FALSE}
model %>% compile(
  optimizer = 'sgd',
  loss = 'mse',
  metrics = list(metric_categorical_crossentropy()))
```

@param name
(Optional) string name of the metric instance.

@param dtype
(Optional) data type of the metric result.

@param from_logits
(Optional) Whether output is expected to be
a logits tensor. By default, we consider that output
encodes a probability distribution.

@param label_smoothing
(Optional) Float in `[0, 1]`.
When > 0, label values are smoothed, meaning the confidence
on label values are relaxed. e.g. `label_smoothing=0.2` means
that we will use a value of 0.1 for label
"0" and 0.9 for label "1".

@param axis
(Optional) Defaults to `-1`.
The dimension along which entropy is computed.

@param y_true
Tensor of one-hot true targets.

@param y_pred
Tensor of predicted targets.

@param ...
Passed on to the Python callable

@export
@family probabilistic metrics
@family metrics
@seealso
+ <https:/keras.io/api/metrics/probabilistic_metrics#categoricalcrossentropy-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/metrics/CategoricalCrossentropy>

