---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Lecun uniform initializer.

@description
Draws samples from a uniform distribution within `[-limit, limit]`, where
`limit = sqrt(3 / fan_in)` (`fan_in` is the number of input units in the
weight tensor).

# Examples
```{r}
# Standalone usage:
initializer <- initializer_lecun_uniform()
values <- initializer(shape = c(2, 2))
```

```{r}
# Usage in a Keras layer:
initializer <- initializer_lecun_uniform()
layer <- layer_dense(units = 3, kernel_initializer = initializer)
```

# Reference
- [Klambauer et al., 2017](https://arxiv.org/abs/1706.02515)

@param seed
An integer or instance of
`random_seed_generator()`.
Used to make the behavior of the initializer
deterministic. Note that an initializer seeded with an integer
or `NULL` (unseeded) will produce the same random values
across multiple calls. To get different random values
across multiple calls, use as seed an instance
of `random_seed_generator()`.

@export
@family initializer
@seealso
+ <https://www.tensorflow.org/api_docs/python/tf/keras/initializers/LecunUniform>
