---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Real-valued Fast Fourier Transform along the last axis of the input.

@description
Computes the 1D Discrete Fourier Transform of a real-valued signal over the
inner-most dimension of input.

Since the Discrete Fourier Transform of a real-valued signal is
Hermitian-symmetric, RFFT only returns the `fft_length / 2 + 1` unique
components of the FFT: the zero-frequency term, followed by the
`fft_length / 2` positive-frequency terms.

Along the axis RFFT is computed on, if `fft_length` is smaller than the
corresponding dimension of the input, the dimension is cropped. If it is
larger, the dimension is padded with zeros.

# Examples
```{r}
x <- k_convert_to_tensor(c(0, 1, 2, 3, 4))
k_rfft(x)
```

```{r}
k_rfft(x, 3)
```

@returns
A list containing two tensors - the real and imaginary parts of the
output.

@param x Input tensor.
@param fft_length An integer representing the number of the fft length. If not
    specified, it is inferred from the length of the last axis of `x`.
    Defaults to `NULL`.

@export
@family ops
@seealso
+ <https:/keras.io/keras_core/api/ops/fft#rfft-function>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/ops/rfft>

