---
title: "Transfer learning and fine-tuning"
author: "[fchollet](https://twitter.com/fchollet), [t-kalinowski](https://github.com/t-kalinowski)"
date: 2021/10/15
description: "Complete guide to transfer learning and fine-tuning in Keras."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transfer learning and fine-tuning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Setup
```{r}
library(tensorflow)
library(keras)
printf <- function(...) writeLines(sprintf(...))
```


## Introduction

**Transfer learning** consists of taking features learned on one problem, and
leveraging them on a new, similar problem. For instance, features from a model that has
learned to identify racoons may be useful to kick-start a model meant to identify skunks.

Transfer learning is usually done for tasks where your dataset has too little data to
 train a full-scale model from scratch.

The most common incarnation of transfer learning in the context of deep learning is the
 following workflow:

1. Take layers from a previously trained model.
2. Freeze them, so as to avoid destroying any of the information they contain during
 future training rounds.
3. Add some new, trainable layers on top of the frozen layers. They will learn to turn
 the old features into predictions on a new dataset.
4. Train the new layers on your dataset.

A last, optional step, is **fine-tuning**, which consists of unfreezing the entire
model you obtained above (or part of it), and re-training it on the new data with a
very low learning rate. This can potentially achieve meaningful improvements, by
 incrementally adapting the pretrained features to the new data.

First, we will go over the Keras `trainable` API in detail, which underlies most
 transfer learning and fine-tuning workflows.

Then, we'll demonstrate the typical workflow by taking a model pretrained on the
ImageNet dataset, and retraining it on the Kaggle "cats vs dogs" classification
 dataset.

This is adapted from
[Deep Learning with R](https://www.manning.com/books/deep-learning-with-r)
 and the 2016 blog post
["building powerful image classification models using very little
 data"](https://blog.keras.io/building-powerful-image-classification-models-using-very-little-data.html).


## Freezing layers: understanding the `trainable` attribute

Layers and models have three weight attributes:

- `weights` is the list of all weights variables of the layer.
- `trainable_weights` is the list of those that are meant to be updated (via gradient
 descent) to minimize the loss during training.
- `non_trainable_weights` is the list of those that aren't meant to be trained.
 Typically they are updated by the model during the forward pass.

**Example: the `Dense` layer has 2 trainable weights (kernel and bias)**

```{r, hold = TRUE}
layer <- layer_dense(units = 3)
layer$build(shape(NULL, 4))

printf("weights: %s", length(layer$weights))
printf("trainable_weights: %s", length(layer$trainable_weights))
printf("non_trainable_weights: %s", length(layer$non_trainable_weights))
```

In general, all weights are trainable weights. The only built-in layer that has
non-trainable weights is `layer_batch_normalization()`. It uses non-trainable weights
 to keep track of the mean and variance of its inputs during training.
To learn how to use non-trainable weights in your own custom layers, see the
[guide to writing new layers from scratch](https://keras.rstudio.com/articles/new-guides/making_new_layers_and_models_via_subclassing.html).

**Example: The layer instance returned by `layer_batch_normalization()` has 2 trainable weights and 2 non-trainable
 weights**
```{r, hold = TRUE}
layer <- layer_batch_normalization()
layer$build(shape(NULL, 4))

printf("weights: %s", length(layer$weights))
printf("trainable_weights: %s", length(layer$trainable_weights))
printf("non_trainable_weights: %s", length(layer$non_trainable_weights))
```

Layers and models also feature a boolean attribute `trainable`. Its value can be changed.
Setting `layer$trainable` to `FALSE` moves all the layer's weights from trainable to
non-trainable.  This is called "freezing" the layer: the state of a frozen layer won't
be updated during training (either when training with `fit()` or when training with
 any custom loop that relies on `trainable_weights` to apply gradient updates).

**Example: setting `trainable` to `False`**
```{r, hold = TRUE}
layer = layer_dense(units = 3)
layer$build(shape(NULL, 4))  # Create the weights
layer$trainable <- FALSE     # Freeze the layer

printf("weights: %s", length(layer$weights))
printf("trainable_weights: %s", length(layer$trainable_weights))
printf("non_trainable_weights: %s", length(layer$non_trainable_weights))
```

When a trainable weight becomes non-trainable, its value is no longer updated during
 training.

```{r}
# Make a model with 2 layers
layer1 <- layer_dense(units = 3, activation = "relu")
layer2 <- layer_dense(units = 3, activation = "sigmoid")
model <- keras_model_sequential(input_shape = c(3)) %>%
  layer1() %>%
  layer2()

# Freeze the first layer
layer1$trainable <- FALSE

# Keep a copy of the weights of layer1 for later reference
initial_layer1_weights_values <- get_weights(layer1)

# Train the model
model %>% compile(optimizer = "adam", loss = "mse")
model %>% fit(k_random_normal(c(2, 3)), k_random_normal(c(2, 3)))

# Check that the weights of layer1 have not changed during training
final_layer1_weights_values <- get_weights(layer1)
stopifnot(all.equal(initial_layer1_weights_values, final_layer1_weights_values))
```

Do not confuse the `layer$trainable` attribute with the `training` argument in a
layer instance's `call` signature `layer(training =)` (which controls whether
the layer should run its forward pass in inference mode or training mode).
For more information, see the [Keras FAQ](
  https://keras.io/getting_started/faq/#whats-the-difference-between-the-training-argument-in-call-and-the-trainable-attribute).

## Recursive setting of the `trainable` attribute

If you set `trainable = FALSE` on a model or on any layer that has sublayers,
all child layers become non-trainable as well.

**Example:**
```{r}
inner_model <- keras_model_sequential(input_shape = c(3)) %>%
  layer_dense(3, activation = "relu") %>%
  layer_dense(3, activation = "relu")

model <- keras_model_sequential(input_shape = c(3)) %>%
  inner_model() %>%
  layer_dense(3, activation = "sigmoid")


model$trainable <- FALSE  # Freeze the outer model

stopifnot(inner_model$trainable == FALSE)             # All layers in `model` are now frozen
stopifnot(inner_model$layers[[1]]$trainable == FALSE)  # `trainable` is propagated recursively
```

## The typical transfer-learning workflow

This leads us to how a typical transfer learning workflow can be implemented in Keras:

1. Instantiate a base model and load pre-trained weights into it.
2. Freeze all layers in the base model by setting `trainable = FALSE`.
3. Create a new model on top of the output of one (or several) layers from the base
 model.
4. Train your new model on your new dataset.

Note that an alternative, more lightweight workflow could also be:

1. Instantiate a base model and load pre-trained weights into it.
2. Run your new dataset through it and record the output of one (or several) layers
 from the base model. This is called **feature extraction**.
3. Use that output as input data for a new, smaller model.

A key advantage of that second workflow is that you only run the base model once on
 your data, rather than once per epoch of training. So it's a lot faster and cheaper.

An issue with that second workflow, though, is that it doesn't allow you to dynamically
modify the input data of your new model during training, which is required when doing
data augmentation, for instance. Transfer learning is typically used for tasks when
your new dataset has too little data to train a full-scale model from scratch, and in
such scenarios data augmentation is very important. So in what follows, we will focus
 on the first workflow.

Here's what the first workflow looks like in Keras:

First, instantiate a base model with pre-trained weights.


```{r}
base_model <- application_xception(
  weights = 'imagenet', # Load weights pre-trained on ImageNet.
  input_shape = c(150, 150, 3),
  include_top = FALSE # Do not include the ImageNet classifier at the top.
)
```

Then, freeze the base model.

```{r}
base_model$trainable <- FALSE
```


Create a new model on top.
```{r}
inputs <- layer_input(c(150, 150, 3))

outputs <- inputs %>%
  # We make sure that the base_model is running in inference mode here,
  # by passing `training=FALSE`. This is important for fine-tuning, as you will
  # learn in a few paragraphs.
  base_model(training=FALSE) %>%

  # Convert features of shape `base_model$output_shape[-1]` to vectors
  layer_global_average_pooling_2d() %>%

  # A Dense classifier with a single unit (binary classification)
  layer_dense(1)

model <- keras_model(inputs, outputs)
```


Train the model on new data.

```{r, eval = FALSE}
model %>%
  compile(optimizer = optimizer_adam(),
          loss = loss_binary_crossentropy(from_logits = TRUE),
          metrics = metric_binary_accuracy()) %>%
  fit(new_dataset, epochs = 20, callbacks = ..., validation_data = ...)
```

## Fine-tuning

Once your model has converged on the new data, you can try to unfreeze all or part of
 the base model and retrain the whole model end-to-end with a very low learning rate.

This is an optional last step that can potentially give you incremental improvements.
 It could also potentially lead to quick overfitting -- keep that in mind.

It is critical to only do this step *after* the model with frozen layers has been
trained to convergence. If you mix randomly-initialized trainable layers with
trainable layers that hold pre-trained features, the randomly-initialized layers will
cause very large gradient updates during training, which will destroy your pre-trained
 features.

It's also critical to use a very low learning rate at this stage, because
you are training a much larger model than in the first round of training, on a dataset
 that is typically very small.
As a result, you are at risk of overfitting very quickly if you apply large weight
 updates. Here, you only want to re-adapt the pretrained weights in an incremental way.

This is how to implement fine-tuning of the whole base model:

```{r, eval = FALSE}
# Unfreeze the base model
base_model$trainable <- TRUE

# It's important to recompile your model after you make any changes
# to the `trainable` attribute of any inner layer, so that your changes
# are taken into account
model %>% compile(
  optimizer = optimizer_adam(1e-5), # Very low learning rate
  loss = loss_binary_crossentropy(from_logits = TRUE),
  metrics = metric_binary_accuracy()
)

# Train end-to-end. Be careful to stop before you overfit!
model %>% fit(new_dataset, epochs=10, callbacks=..., validation_data=...)
```


**Important note about `compile()` and `trainable`**

Calling `compile()` on a model is meant to "freeze" the behavior of that model. This
 implies that the `trainable`
attribute values at the time the model is compiled should be preserved throughout the
 lifetime of that model,
until `compile` is called again. Hence, if you change any `trainable` value, make sure
 to call `compile()` again on your
model for your changes to be taken into account.

**Important notes about `layer_batch_normalization()`**

Many image models contain `BatchNormalization` layers. That layer is a special case on
 every imaginable count. Here are a few things to keep in mind.

- `BatchNormalization` contains 2 non-trainable weights that get updated during
training. These are the variables tracking the mean and variance of the inputs.
- When you set `bn_layer$trainable = FALSE`, the `BatchNormalization` layer will
run in inference mode, and will not update its mean and variance statistics. This is not
the case for other layers in general, as
[weight trainability and inference/training modes are two orthogonal concepts](
  https://keras.io/getting_started/faq/#whats-the-difference-between-the-training-argument-in-call-and-the-trainable-attribute).
But the two are tied in the case of the `BatchNormalization` layer.
- When you unfreeze a model that contains `BatchNormalization` layers in order to do
fine-tuning, you should keep the `BatchNormalization` layers in inference mode by
 passing `training = FALSE` when calling the base model.
Otherwise the updates applied to the non-trainable weights will suddenly destroy
what the model has learned.

You'll see this pattern in action in the end-to-end example at the end of this guide.


## Transfer learning and fine-tuning with a custom training loop

If instead of `fit()`, you are using your own low-level training loop, the workflow
stays essentially the same. You should be careful to only take into account the list
 `model$trainable_weights` when applying gradient updates:

```{r, eval = FALSE}
# Create base model
base_model = application_xception(
  weights = 'imagenet',
  input_shape = c(150, 150, 3),
  include_top = FALSE
)

# Freeze base model
base_model$trainable = FALSE

# Create new model on top.
inputs <- layer_input(shape = c(150, 150, 3))
outputs <- inputs %>%
  base_model(training = FALSE) %>%
  layer_global_average_pooling_2d() %>%
  layer_dense(1)
model <- keras_model(inputs, outputs)

loss_fn <- loss_binary_crossentropy(from_logits = TRUE)
optimizer <- optimizer_adam()

# helper to zip gradients with weights
xyz <- function(...) .mapply(c, list(...), NULL)

# Iterate over the batches of a dataset.
library(tfdatasets)
new_dataset <- ...

while(!is.null(batch <- iter_next(new_dataset))) {
  c(inputs, targets) %<-% batch
  # Open a GradientTape.
  with(tf$GradientTape() %as% tape, {
    # Forward pass.
    predictions = model(inputs)
    # Compute the loss value for this batch.
    loss_value = loss_fn(targets, predictions)
  })
  # Get gradients of loss w.r.t. the *trainable* weights.
  gradients <- tape$gradient(loss_value, model$trainable_weights)
  # Update the weights of the model.
  optimizer$apply_gradients(xyz(gradients, model$trainable_weights))
}
```


Likewise for fine-tuning.

## An end-to-end example: fine-tuning an image classification model on a cats vs. dogs dataset

To solidify these concepts, let's walk you through a concrete end-to-end transfer
learning and fine-tuning example. We will load the Xception model, pre-trained on
 ImageNet, and use it on the Kaggle "cats vs. dogs" classification dataset.

### Getting the data

First, let's fetch the cats vs. dogs dataset using TFDS. If you have your own dataset,
you'll probably want to use the utility
`image_dataset_from_directory()` to generate similar labeled
 dataset objects from a set of images on disk filed into class-specific folders.

Transfer learning is most useful when working with very small datasets. To keep our
dataset small, we will use 40% of the original training data (25,000 images) for
 training, 10% for validation, and 10% for testing.

```{r}
# reticulate::py_install("tensorflow_datasets", pip = TRUE)
tfds <- reticulate::import("tensorflow_datasets")

c(train_ds, validation_ds, test_ds) %<-% tfds$load(
    "cats_vs_dogs",
    # Reserve 10% for validation and 10% for test
    split = c("train[:40%]", "train[40%:50%]", "train[50%:60%]"),
    as_supervised=TRUE  # Include labels
)

printf("Number of training samples: %d", length(train_ds))
printf("Number of validation samples: %d", length(validation_ds) )
printf("Number of test samples: %d", length(test_ds))
```

These are the first 9 images in the training dataset -- as you can see, they're all
 different sizes.
```{r}
library(tfdatasets)

par(mfrow = c(3, 3), mar = c(1,0,1.5,0))
train_ds %>%
  dataset_take(9) %>%
  as_array_iterator() %>%
  iterate(function(batch) {
    c(image, label) %<-% batch
    plot(as.raster(image, max = 255))
    title(sprintf("label: %s   size: %s",
                  label, paste(dim(image), collapse = " x ")))
  })
```

We can also see that label 1 is "dog" and label 0 is "cat".

### Standardizing the data

Our raw images have a variety of sizes. In addition, each pixel consists of 3 integer
values between 0 and 255 (RGB level values). This isn't a great fit for feeding a
 neural network. We need to do 2 things:

- Standardize to a fixed image size. We pick 150x150.
- Normalize pixel values between -1 and 1. We'll do this using a `layer_normalization()` as
 part of the model itself.

In general, it's a good practice to develop models that take raw data as input, as
opposed to models that take already-preprocessed data. The reason being that, if your
model expects preprocessed data, any time you export your model to use it elsewhere
(in a web browser, in a mobile app), you'll need to reimplement the exact same
preprocessing pipeline. This gets very tricky very quickly. So we should do the least
 possible amount of preprocessing before hitting the model.

Here, we'll do image resizing in the data pipeline (because a deep neural network can
only process contiguous batches of data), and we'll do the input value scaling as part
 of the model, when we create it.

Let's resize images to 150x150:

```{r}
library(magrittr, include.only = "%<>%")
size <- as.integer(c(150, 150))
train_ds      %<>% dataset_map(function(x, y) list(tf$image$resize(x, size), y))
validation_ds %<>% dataset_map(function(x, y) list(tf$image$resize(x, size), y))
test_ds       %<>% dataset_map(function(x, y) list(tf$image$resize(x, size), y))
```

Besides, let's batch the data and use caching and prefetching to optimize loading speed.
```{r}
dataset_cache_batch_prefetch <- function(dataset, batch_size = 32, buffer_size = 10) {
  dataset %>%
    dataset_cache() %>%
    dataset_batch(batch_size) %>%
    dataset_prefetch(buffer_size)
}

train_ds      %<>% dataset_cache_batch_prefetch()
validation_ds %<>% dataset_cache_batch_prefetch()
test_ds       %<>% dataset_cache_batch_prefetch()
```

### Using random data augmentation

When you don't have a large image dataset, it's a good practice to artificially
 introduce sample diversity by applying random yet realistic transformations to
the training images, such as random horizontal flipping or small random rotations. This
helps expose the model to different aspects of the training data while slowing down
 overfitting.

```{r}
data_augmentation <- keras_model_sequential() %>%
  layer_random_flip("horizontal") %>%
  layer_random_rotation(.1)
```

Let's visualize what the first image of the first batch looks like after various random
 transformations:

```{r}
batch <- train_ds %>%
  dataset_take(1) %>%
  as_iterator() %>% iter_next()

c(images, labels) %<-% batch
first_image <- images[1, all_dims(), drop = FALSE]
augmented_image <- data_augmentation(first_image, training = TRUE)

plot_image <- function(image, main = deparse1(substitute(image))) {
  image %>%
    k_squeeze(1) %>% # drop batch dim
    as.array() %>%   # convert from tensor to R array
    as.raster(max = 255) %>%
    plot()

  if(!is.null(main))
    title(main)
}

par(mfrow = c(2, 2), mar = c(1, 1, 1.5, 1))
plot_image(first_image)
plot_image(augmented_image)
plot_image(data_augmentation(first_image, training = TRUE), "augmented 2")
plot_image(data_augmentation(first_image, training = TRUE), "augmented 3")
```




## Build a model

Now let's build a model that follows the blueprint we've explained earlier.

Note that:

- We add `layer_rescaling()` to scale input values (initially in the `[0, 255]`
 range) to the `[-1, 1]` range.
- We add a `layer_dropout()` before the classification layer, for regularization.
- We make sure to pass `training = FALSE` when calling the base model, so that
it runs in inference mode, so that batchnorm statistics don't get updated
even after we unfreeze the base model for fine-tuning.

```{r}
base_model = application_xception(
  weights = "imagenet", # Load weights pre-trained on ImageNet.
  input_shape = c(150, 150, 3),
  include_top = FALSE # Do not include the ImageNet classifier at the top.
)

# Freeze the base_model
base_model$trainable <- FALSE

# Create new model on top
inputs = layer_input(shape = c(150, 150, 3))

outputs <- inputs %>%
  data_augmentation() %>%   # Apply random data augmentation

  # Pre-trained Xception weights requires that input be scaled
  # from (0, 255) to a range of (-1., +1.), the rescaling layer
  # outputs: `(inputs * scale) + offset`
  layer_rescaling(scale = 1 / 127.5, offset = -1) %>%

  # The base model contains batchnorm layers. We want to keep them in inference mode
  # when we unfreeze the base model for fine-tuning, so we make sure that the
  # base_model is running in inference mode here.
  base_model(training = FALSE) %>%
  layer_global_average_pooling_2d() %>%
  layer_dropout(.2) %>%
  layer_dense(1)

model <- keras_model(inputs, outputs)
model
```

## Train the top layer
```{r}
model %>% compile(
  optimizer = optimizer_adam(),
  loss = loss_binary_crossentropy(from_logits = TRUE),
  metrics = metric_binary_accuracy()
)

epochs <- 20
model %>% fit(train_ds, epochs = epochs, validation_data = validation_ds)
```

## Do a round of fine-tuning of the entire model

Finally, let's unfreeze the base model and train the entire model end-to-end with a low
 learning rate.

Importantly, although the base model becomes trainable, it is still running in
inference mode since we passed `training = FALSE` when calling it when we built the
model. This means that the batch normalization layers inside won't update their batch
statistics. If they did, they would wreck havoc on the representations learned by the
 model so far.

```{r}
# Unfreeze the base_model. Note that it keeps running in inference mode
# since we passed `training = FALSE` when calling it. This means that
# the batchnorm layers will not update their batch statistics.
# This prevents the batchnorm layers from undoing all the training
# we've done so far.
base_model$trainable <- TRUE
model

model %>% compile(
  optimizer = optimizer_adam(1e-5),
  loss = loss_binary_crossentropy(from_logits = TRUE),
  metrics = metric_binary_accuracy()
)

epochs <- 10
model %>% fit(train_ds, epochs = epochs, validation_data = validation_ds)
```

After 10 epochs, fine-tuning gains us a nice improvement here.
