---
title: Introduction to Keras for engineers
author: '[fchollet](https://twitter.com/fchollet)'
date-created: 2023/07/10
last-modified: 2023/07/10
description: First contact with Keras 3.
accelerator: GPU
output: rmarkdown::html_vignette
knit: ({source(here::here("tools/knit.R")); knit_vignette})
tether: https://raw.githubusercontent.com/keras-team/keras-io/master/guides/intro_to_keras_for_engineers.py
editor_options:
  chunk_output_type: inline
---

## Introduction

Keras 3 is a deep learning framework
works with TensorFlow, JAX, and PyTorch interchangeably.
This notebook will walk you through key Keras 3 workflows.

Let's start by installing Keras 3:

pip install keras --upgrade --quiet

## Setup

We're going to be using the tensorflow backend here -- but you can
edit the string below to `"jax"` or `"torch"` and hit
"Restart runtime", and the whole notebook will run just the same!
This entire guide is backend-agnostic.

```{r}
library(tensorflow, exclude = c("shape", "set_random_seed"))
library(keras3)

# Note that you must configure the backend
# before calling any other keras functions.
# The backend cannot be changed once the
# package is imported.
use_backend("tensorflow")
```

## A first example: A MNIST convnet

Let's start with the Hello World of ML: training a convnet
to classify MNIST digits.

Here's the data:

```{r}
# Load the data and split it between train and test sets
c(c(x_train, y_train), c(x_test, y_test)) %<-% keras3::dataset_mnist()

# Scale images to the [0, 1] range
x_train <- x_train / 255
x_test <- x_test / 255

# Make sure images have shape (28, 28, 1)
x_train <- array_reshape(x_train, c(-1, 28, 28, 1))
x_test <- array_reshape(x_test, c(-1, 28, 28, 1))

dim(x_train)
dim(x_test)
```

Here's our model.

Different model-building options that Keras offers include:

- [The Sequential API](sequential_model.html) (what we use below)
- [The Functional API](functional_api.html) (most typical)
- [Writing your own models yourself via subclassing](making_new_layers_and_models_via_subclassing.html) (for advanced use cases)

```{r}
# Model parameters
num_classes <- 10
input_shape <- c(28, 28, 1)

model <- keras_model_sequential(input_shape = input_shape)
model |>
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") |>
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") |>
  layer_max_pooling_2d(pool_size = c(2, 2)) |>
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") |>
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") |>
  layer_global_average_pooling_2d() |>
  layer_dropout(rate = 0.5) |>
  layer_dense(units = num_classes, activation = "softmax")
```

Here's our model summary:

```{r}
summary(model)
```

We use the `compile()` method to specify the optimizer, loss function,
and the metrics to monitor. Note that with the JAX and TensorFlow backends,
XLA compilation is turned on by default.

```{r}
model |> compile(
  optimizer = "adam",
  loss = "sparse_categorical_crossentropy",
  metrics = list(
    metric_sparse_categorical_accuracy(name = "acc")
  )
)
```

Let's train and evaluate the model. We'll set aside a validation split of 15%
of the data during training to monitor generalization on unseen data.

```{r}
batch_size <- 128
epochs <- 10

callbacks <- list(
  callback_model_checkpoint(filepath="model_at_epoch_{epoch}.keras"),
  callback_early_stopping(monitor="val_loss", patience=2)
)

model |> fit(
  x_train, y_train,
  batch_size = batch_size,
  epochs = epochs,
  validation_split = 0.15,
  callbacks = callbacks
)
score <- model |> evaluate(x_test, y_test, verbose = 0)
```

During training, we were saving a model at the end of each epoch. You
can also save the model in its latest state like this:

```{r}
save_model(model, "final_model.keras", overwrite=TRUE)
```

And reload it like this:

```{r}
model <- load_model("final_model.keras")
```
```{r, include = FALSE}
unlink("final_model.keras")
```

Next, you can query predictions of class probabilities with `predict()`:

```{r}
predictions <- model |> predict(x_test)
dim(predictions)
```

That's it for the basics!

## Writing cross-framework custom components

Keras enables you to write custom Layers, Models, Metrics, Losses, and Optimizers
that work across TensorFlow, JAX, and PyTorch with the same codebase. Let's take a look
at custom layers first.

The `op_` namespace contains:

- An implementation of the NumPy API, e.g. `op_stack` or `op_matmul`.
- A set of neural network specific ops that are absent from NumPy, such as `op_conv`
or `op_binary_crossentropy`.

Let's make a custom `Dense` layer that works with all backends:

```{r}
layer_my_dense <- Layer(
  classname = "MyDense",
  initialize = function(units, activation = NULL, name = NULL, ...) {
    super$initialize(name = name, ...)
    self$units <- units
    self$activation <- activation
  },
  build = function(input_shape) {
    input_dim <- tail(input_shape, 1)
    self$w <- self$add_weight(
      shape = shape(input_dim, self$units),
      initializer = initializer_glorot_normal(),
      name = "kernel",
      trainable = TRUE
    )
    self$b <- self$add_weight(
      shape = shape(self$units),
      initializer = initializer_zeros(),
      name = "bias",
      trainable = TRUE
    )
  },
  call = function(inputs) {
    # Use Keras ops to create backend-agnostic layers/metrics/etc.
    x <- op_matmul(inputs, self$w) + self$b
    if (!is.null(self$activation))
      x <- self$activation(x)
    x
  }
)
```

Next, let's make a custom `Dropout` layer that relies on the `random_*`
namespace:

```{r}
layer_my_dropout <- Layer(
  "MyDropout",
  initialize = function(rate, name = NULL, seed = NULL, ...) {
    super$initialize(name = name)
    self$rate <- rate
    # Use seed_generator for managing RNG state.
    # It is a state element and its seed variable is
    # tracked as part of `layer$variables`.
    self$seed_generator <- random_seed_generator(seed)
  },
  call = function(inputs) {
    # Use `keras3::random_*` for random ops.
    random_dropout(inputs, self$rate, seed = self$seed_generator)
  }
)
```

Next, let's write a custom subclassed model that uses our two custom layers:

```{r}
MyModel <- Model(
  "MyModel",
  initialize = function(num_classes, ...) {
    super$initialize(...)
    self$conv_base <-
      keras_model_sequential() |>
      layer_conv_2d(64, kernel_size = c(3, 3), activation = "relu") |>
      layer_conv_2d(64, kernel_size = c(3, 3), activation = "relu") |>
      layer_max_pooling_2d(pool_size = c(2, 2)) |>
      layer_conv_2d(128, kernel_size = c(3, 3), activation = "relu") |>
      layer_conv_2d(128, kernel_size = c(3, 3), activation = "relu") |>
      layer_global_average_pooling_2d()

    self$dp <- layer_my_dropout(rate = 0.5)
    self$dense <- layer_my_dense(units = num_classes,
                                 activation = activation_softmax)
  },
  call = function(inputs) {
    inputs |>
      self$conv_base() |>
      self$dp() |>
      self$dense()
  }
)
```

Let's compile it and fit it:

```{r}
model <- MyModel(num_classes = 10)
model |> compile(
  loss = loss_sparse_categorical_crossentropy(),
  optimizer = optimizer_adam(learning_rate = 1e-3),
  metrics = list(
    metric_sparse_categorical_accuracy(name = "acc")
  )
)

model |> fit(
  x_train, y_train,
  batch_size = batch_size,
  epochs = 1, # For speed
  validation_split = 0.15
)
```

## Training models on arbitrary data sources

All Keras models can be trained and evaluated on a wide variety of data sources,
independently of the backend you're using. This includes:

- Arrays
- Dataframes
- TensorFlow `tf_dataset` objects
- PyTorch `DataLoader` objects
- Keras `PyDataset` objects

They all work whether you're using TensorFlow, JAX, or PyTorch as your Keras backend.

<!-- Let's try it out with PyTorch `DataLoaders`: -->

<!-- ```python -->
<!-- import torch -->

<!-- # Create a TensorDataset -->
<!-- train_torch_dataset = torch.utils.data.TensorDataset( -->
<!--     torch.from_numpy(x_train), torch.from_numpy(y_train) -->
<!-- ) -->
<!-- val_torch_dataset = torch.utils.data.TensorDataset( -->
<!--     torch.from_numpy(x_test), torch.from_numpy(y_test) -->
<!-- ) -->

<!-- # Create a DataLoader -->
<!-- train_dataloader = torch.utils.data.DataLoader( -->
<!--     train_torch_dataset, batch_size=batch_size, shuffle=True -->
<!-- ) -->
<!-- val_dataloader = torch.utils.data.DataLoader( -->
<!--     val_torch_dataset, batch_size=batch_size, shuffle=False -->
<!-- ) -->

<!-- model = MyModel(num_classes=10) -->
<!-- model.compile( -->
<!--     loss=keras.losses.SparseCategoricalCrossentropy(), -->
<!--     optimizer=keras.optimizers.Adam(learning_rate=1e-3), -->
<!--     metrics=[ -->
<!--         keras.metrics.SparseCategoricalAccuracy(name="acc"), -->
<!--     ], -->
<!-- ) -->
<!-- model.fit(train_dataloader, epochs=1, validation_data=val_dataloader) -->
<!-- ``` -->

Let's try this out with `tf_dataset`:

```{r}
library(tfdatasets, exclude = "shape")

train_dataset <- list(x_train, y_train) |>
  tensor_slices_dataset() |>
  dataset_batch(batch_size) |>
  dataset_prefetch(buffer_size = tf$data$AUTOTUNE)

test_dataset <- list(x_test, y_test) |>
  tensor_slices_dataset() |>
  dataset_batch(batch_size) |>
  dataset_prefetch(buffer_size = tf$data$AUTOTUNE)

model <- MyModel(num_classes = 10)
model |> compile(
  loss = loss_sparse_categorical_crossentropy(),
  optimizer = optimizer_adam(learning_rate = 1e-3),
  metrics = list(
    metric_sparse_categorical_accuracy(name = "acc")
  )
)

model |> fit(train_dataset, epochs = 1, validation_data = test_dataset)
```

## Further reading

This concludes our short overview of the new multi-backend capabilities
of Keras 3. Next, you can learn about:

### How to customize what happens in `fit()`

Want to implement a non-standard training algorithm yourself but still want to benefit from
the power and usability of `fit()`? It's easy to customize
`fit()` to support arbitrary use cases:

- [Customizing what happens in `fit()` with TensorFlow](custom_train_step_in_tensorflow.html)
<!-- - [Customizing what happens in `fit()` with JAX](https://keras.io/guides/custom_train_step_in_jax/) -->
<!-- - [Customizing what happens in `fit()` with PyTorch](https://keras.io/guides/custom_train_step_in_pytorch/) -->

## How to write custom training loops

- [Writing a training loop from scratch in TensorFlow](writing_a_custom_training_loop_in_tensorflow.html)
<!-- - [Writing a training loop from scratch in JAX](https://keras.io/guides/writing_a_custom_training_loop_in_jax/) -->
<!-- - [Writing a training loop from scratch in PyTorch](https://keras.io/guides/writing_a_custom_training_loop_in_torch/) -->

## How to distribute training

- [Guide to distributed training with TensorFlow](distributed_training_with_tensorflow.html)
<!-- - [JAX distributed training example](https://github.com/keras-team/keras/blob/master/examples/demo_jax_distributed.py) -->
<!-- - [PyTorch distributed training example](https://github.com/keras-team/keras/blob/master/examples/demo_torch_multi_gpu.py) -->

Enjoy the library! 🚀
