---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Concatenates a list of inputs.

@description
It takes as input a list of tensors, all of the same shape except
for the concatenation axis, and returns a single tensor that is the
concatenation of all inputs.

# Examples
```{r}
x <- k_arange(20) |> k_reshape(c(2, 2, 5))
y <- k_arange(20, 40) |> k_reshape(c(2, 2, 5))
layer_concatenate(x, y, axis = 2)
```
Usage in a Keras model:

```{r}
x1 <- k_arange(10)     |> k_reshape(c(5, 2)) |> layer_dense(8)
x2 <- k_arange(10, 20) |> k_reshape(c(5, 2)) |> layer_dense(8)
y <- layer_concatenate(x1, x2)
```

@returns
    A tensor, the concatenation of the inputs alongside axis `axis`.

@param axis
Axis along which to concatenate.

@param ...
Standard layer keyword arguments.

@param inputs
layers to combine

@export
@family concatenate merging layers
@family merging layers
@family layers
@seealso
+ <https:/keras.io/api/layers/merging_layers/concatenate#concatenate-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/layers/Concatenate>
