---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes mean absolute percentage error between `y_true` and `y_pred`.

@description
Formula:

```{r, eval = FALSE}
loss <- 100 * mean(abs((y_true - y_pred) / y_true), axis=-1)
```

Division by zero is prevented by dividing by `maximum(y_true, epsilon)`
where `epsilon = keras$backend$epsilon()`
(default to `1e-7`).

# Examples
Standalone usage:

```{r}
m <- metric_mean_absolute_percentage_error()
m$update_state(rbind(c(0, 1), c(0, 0)), rbind(c(1, 1), c(0, 0)))
m$result()

m$reset_state()
m$update_state(rbind(c(0, 1), c(0, 0)), rbind(c(1, 1), c(0, 0)),
               sample_weight = c(1, 0))
m$result()
```

Usage with `compile()` API:

```{r, eval = FALSE}
model %>% compile(
    optimizer = 'sgd',
    loss = 'mse',
    metrics = list(metric_mean_absolute_percentage_error()))
```

@param name
(Optional) string name of the metric instance.

@param dtype
(Optional) data type of the metric result.

@param y_true
Ground truth values with shape = `[batch_size, d0, .. dN]`.

@param y_pred
The predicted values with shape = `[batch_size, d0, .. dN]`.

@param ...
Passed on to the Python callable

@export
@family regression metrics
@family metrics
@seealso
+ <https:/keras.io/api/metrics/regression_metrics#meanabsolutepercentageerror-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/metrics/MeanAbsolutePercentageError>

