---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
Computes the mean of absolute difference between labels and predictions.

@description
Formula:

```{r, eval = FALSE}
loss <- mean(abs(y_true - y_pred))
```

# Examples
```{r}
y_true <- random_uniform(c(2, 3), 0, 2)
y_pred <- random_uniform(c(2, 3))
loss <- loss_mean_absolute_error(y_true, y_pred)
```

@returns
Mean absolute error values with shape = `[batch_size, d0, .. dN-1]`.

@param reduction
Type of reduction to apply to the loss. In almost all cases
this should be `"sum_over_batch_size"`.
Supported options are `"sum"`, `"sum_over_batch_size"` or `NULL`.

@param name
Optional name for the loss instance.

@param y_true
Ground truth values with shape = `[batch_size, d0, .. dN]`.

@param y_pred
The predicted values with shape = `[batch_size, d0, .. dN]`.

@param ...
Passed on to the Python callable

@export
@family losses
@seealso
+ <https:/keras.io/api/losses/regression_losses#meanabsoluteerror-class>
+ <https://www.tensorflow.org/api_docs/python/tf/keras/losses/MeanAbsoluteError>

