---
knit: ({source(here::here("tools/knit.R")); knit_man_src})
---
A layer that uses `einsum` as the backing computation.

@description
This layer can perform einsum calculations of arbitrary dimensionality.

# Examples
**Biased dense layer with einsums**

This example shows how to instantiate a standard Keras dense layer using
einsum operations. This example is equivalent to
`layer_Dense(64, use_bias=TRUE)`.

```{r}
input <- layer_input(shape = c(32))
output <- input |>
  layer_einsum_dense("ab,bc->ac",
                     output_shape = 64,
                     bias_axes = "c")
output # shape: (NULL, 64)
```

**Applying a dense layer to a sequence**

This example shows how to instantiate a layer that applies the same dense
operation to every element in a sequence. Here, the `output_shape` has two
values (since there are two non-batch dimensions in the output); the first
dimension in the `output_shape` is `NA`, because the sequence dimension
`b` has an unknown shape.

```{r}
input <- layer_input(shape = c(32, 128))
output <- input |>
  layer_einsum_dense("abc,cd->abd",
                     output_shape = c(NA, 64),
                     bias_axes = "d")
output  #shape (NA, 32, 64)
```

**Applying a dense layer to a sequence using ellipses**

This example shows how to instantiate a layer that applies the same dense
operation to every element in a sequence, but uses the ellipsis notation
instead of specifying the batch and sequence dimensions.

Because we are using ellipsis notation and have specified only one axis, the
`output_shape` arg is a single value. When instantiated in this way, the
layer can handle any number of sequence dimensions - including the case
where no sequence dimension exists.

```{r}
input <- layer_input(shape = c(32, 128))
output <- input |>
  layer_einsum_dense("...x,xy->...y",
                     output_shape = 64,
                     bias_axes = "y")

output  #shape (NA, 32, 64)
```

@param equation
An equation describing the einsum to perform.
This equation must be a valid einsum string of the form
`ab,bc->ac`, `...ab,bc->...ac`, or
`ab...,bc->ac...` where 'ab', 'bc', and 'ac' can be any valid einsum
axis expression sequence.

@param output_shape
The expected shape of the output tensor
(excluding the batch dimension and any dimensions
represented by ellipses). You can specify `NA` or `NULL` for any dimension
that is unknown or can be inferred from the input shape.

@param activation
Activation function to use. If you don't specify anything,
no activation is applied
(that is, a "linear" activation: `a(x) = x`).

@param bias_axes
A string containing the output dimension(s)
to apply a bias to. Each character in the `bias_axes` string
should correspond to a character in the output portion
of the `equation` string.

@param kernel_initializer
Initializer for the `kernel` weights matrix.

@param bias_initializer
Initializer for the bias vector.

@param kernel_regularizer
Regularizer function applied to the `kernel` weights
matrix.

@param bias_regularizer
Regularizer function applied to the bias vector.

@param kernel_constraint
Constraint function applied to the `kernel` weights
matrix.

@param bias_constraint
Constraint function applied to the bias vector.

@param ...
Base layer keyword arguments, such as `name` and `dtype`.

@param object
Object to compose the layer with. A tensor, array, or sequential model.

@export
@family core layers
@family layers
@seealso
+ <https://www.tensorflow.org/api_docs/python/tf/keras/layers/EinsumDense>

